---
title: "Importing data from flat files with utils"
author: "Mohamad Osman"
date: '2022-06-21'
output:
  pdf_document: default
  html_document: default
---

### **`01-read.csv`**

-   Use `read.csv()` to import `"swimming_pools.csv"` as a data frame with the name `pools`.

-   Print the structure of `pools` using `str()`.

```{r}
swimming_pools_path <- file.path("..","00_Datasets","swimming_pools.csv")
swimming_pools_path
```

```{r}
# Import swimming_pools.csv: pools
pools <- read.csv(swimming_pools_path)

# Print the structure of pools
str(pools)
```

### **`02-stringsAsFactors`**

-   Use `read.csv()` to import the data in `"swimming_pools.csv"` as a data frame called `pools`; make sure that strings are imported as characters, not as factors.

-   Using `str()`, display the structure of the dataset and check that you indeed get character vectors instead of factors.

```{r}
# Import swimming_pools.csv correctly: pools
pools <- read.csv(swimming_pools_path, stringsAsFactors = FALSE)

# Check the structure of pools
str(pools)
```

### **`03-read.delim`**

-   Import the data in `"hotdogs.txt"` with `read.delim()`. Call the resulting data frame `hotdogs`. The variable names are **not** on the first line, so make sure to set the `header` argument appropriately.

-   Call `summary()` on `hotdogs`. This will print out some summary statistics about all variables in the data frame.

```{r}

hotdogs_path <- file.path("..","00_Datasets","hotdogs.txt")

# Import hotdogs.txt: hotdogs
hotdogs <- read.delim(hotdogs_path, sep = "\t", header = FALSE,
                      stringsAsFactors = FALSE)

# Summarize hotdogs
summary(hotdogs)
```

### **`04-read.table`**

-   Finish the `read.table()` call that's been prepared for you. Use the `path` variable, and make sure to set `sep` correctly.

-   Call `head()` on `hotdogs`; this will print the first 6 observations in the data frame.

```{r}
# Path to the hotdogs.txt file: path
hotdogs_path <- file.path("..", "00_Datasets", "hotdogs.txt")

# Import the hotdogs.txt file: hotdogs
hotdogs <- read.table(hotdogs_path, 
                      sep = '\t', 
                      col.names = c("type", "calories", "sodium"))

# Call head() on hotdogs
head(hotdogs)
```

### **`05-Arguments`**

-   Finish the `read.delim()` call to import the data in `"hotdogs.txt"`. It's a tab-delimited file without names in the first row.

-   The code that selects the observation with the lowest calorie count and stores it in the variable `lily` is already available. It uses the function `which.min()`, that returns the index the smallest value in a vector.

-   Do a similar thing for Tom: select the observation with the *most sodium* and store it in `tom`. Use `which.max()` this time.

-   Finally, print both the observations `lily` and `tom`.

```{r}
# Finish the read.delim() call
hotdogs <- read.delim(hotdogs_path, header = FALSE, col.names = c("type", "calories", "sodium"))

# Select the hot dog with the least calories: lily
lily <- hotdogs[which.min(hotdogs$calories), ]

# Select the observation with the most sodium: tom
tom <- hotdogs[which.max(hotdogs$sodium),]

# Print lily and tom
print(lily)
print(tom)

```

### **`06-Column classes`**

-   The `read.delim()` call from before is already included and creates the `hotdogs` data frame. Go ahead and display the structure of `hotdogs`.

-   **Edit** the second `read.delim()` call. Assign the correct vector to the `colClasses` argument. `NA` should be replaced with a character vector: `c("factor", "NULL", "numeric")`.

-   Display the structure of `hotdogs2` and look for the difference.

```{r}
# Previous call to import hotdogs.txt
hotdogs <- read.delim(hotdogs_path, header = FALSE, col.names = c("type", "calories", "sodium"))

# Display structure of hotdogs
str(hotdogs)

# Edit the colClasses argument to import the data correctly: hotdogs2
hotdogs2 <- read.delim(hotdogs_path, header = FALSE, 
                       col.names = c("type", "calories", "sodium"),
                       colClasses = c("factor", "NULL", "numeric"))


# Display structure of hotdogs2
str(hotdogs2)
```

```{r}
head(hotdogs,3)
head(hotdogs2,3)
```

### `The END`
