---
title: "Reproducible Excel work with XLConnect"
author: "Mohamad Osman"
date: '2022-06-22'
output: html_document
---

# **Section 04: Reproducible Excel work with XLConnect**

### **`01-Connect to a workbook`**

-   Load the `XLConnect` package using `library()`; it is already installed on DataCamp's servers.

-   Use `loadWorkbook()` to build a connection to the `"urbanpop.xlsx"` file in R. Call the workbook `my_book`.

-   Print out the class of `my_book`. What does this tell you?

```{r}
# urbanpop.xlsx is available in your working directory
urbanpop_path <- file.path("..", "00_Datasets", "urbanpop.xlsx")

# Load the XLConnect package
# > install.packages("XLConnect", dependencies=TRUE)
library(XLConnect)

# Build connection to urbanpop.xlsx: my_book
my_book <- XLConnect::loadWorkbook(urbanpop_path)

# Print out the class of my_book
class(my_book)
```

### **`02-List and read Excel sheets`**

-   Print out the sheets of the Excel file that `my_book` links to.

-   Import the second sheet in `my_book` as a data frame. Print it out.

```{r}
# XLConnect is already available

# Build connection to urbanpop.xlsx
my_book <- loadWorkbook(urbanpop_path)

# List the sheets in my_book
getSheets(my_book)

# Import the second sheet in my_book
readWorksheet(my_book, sheet = 2 )


```

### **`03-Customize readWorksheet`**

-   Extend the `readWorksheet()` command with the `startCol` and `endCol` arguments to only import the columns 3, 4, and 5 of the second sheet.

-   `urbanpop_sel` no longer contains information about the countries now. Can you write another `readWorksheet()` command that imports only the first column from the second sheet? Store the resulting data frame as `countries`.

-   Use `cbind()` to paste together `countries` and `urbanpop_sel`, in this order. Store the result as `selection`.

```{r}
# XLConnect is already available

# Build connection to urbanpop.xlsx
my_book <- loadWorkbook(urbanpop_path)

# Import columns 3, 4, and 5 from second sheet in my_book: urbanpop_sel
urbanpop_sel <- readWorksheet(my_book, sheet = 2,
                              startCol = 3 , endCol = 5)

# Import first column from second sheet in my_book: countries
countries <- readWorksheet(my_book, sheet = 2,
                           startCol = 1 , endCol = 1)

# cbind() urbanpop_sel and countries together: selection
selection <- cbind(countries, urbanpop_sel)
selection
```

### **`04-Add worksheet`**

-   Use `createSheet()`, to create a new sheet in `my_book`, named `"data_summary"`.

-   Use `getSheets()` to verify that `my_book` now represents an Excel file with four sheets.

```{r}
# XLConnect is already available

# Build connection to urbanpop.xlsx
my_book <- loadWorkbook(urbanpop_path)

# Add a worksheet to my_book, named "data_summary"
createSheet(my_book, name = "data_summary")

# Use getSheets() on my_book
getSheets(my_book)
```

### **`05-Populate worksheet`**

-   Use `writeWorksheet()` to populate the `"data_summary"` sheet with the `summ` data frame.

-   Call `saveWorkbook()` to store the adapted Excel workbook as a new file, `"summary.xlsx"`.

```{r}
# XLConnect is already available
in_xlsx_file <- file.path("..", "00_Datasets", "urbanpop.xlsx")
out_xlsx_file <- file.path("..", "00_Datasets", "summary.xlsx")

# Build connection to urbanpop.xlsx
my_book <- loadWorkbook(in_xlsx_file)

# Add a worksheet to my_book, named "data_summary"
createSheet(my_book, "data_summary")

# Create data frame: summ
sheets <- getSheets(my_book)[1:3]
dims <- sapply(sheets, function(x) dim(readWorksheet(my_book, sheet = x)), USE.NAMES = FALSE)
summ <- data.frame(sheets = sheets,
                   nrows = dims[1, ],
                   ncols = dims[2, ])

# Add data in summ to "data_summary" sheet
writeWorksheet(my_book, summ, sheet = "data_summary")

# Save workbook as summary.xlsx
saveWorkbook(my_book, out_xlsx_file)
```

### **`06-Renaming sheets`**

-   Use `renameSheet()` to rename the fourth sheet to `"summary"`.

-   Next, call `getSheets()` on `my_book` to print out the sheet names.

-   Finally, make sure to actually save the `my_book` object to a new Excel file, `"renamed.xlsx"`.

```{r}
# XLConnect is already available
in_xlsx_file <- file.path("..", "00_Datasets", "urbanpop.xlsx")
out_xlsx_file <- file.path("..", "00_Datasets", "urbanpop_summary.xlsx")

# Build connection to urbanpop.xlsx
my_book <- loadWorkbook(in_xlsx_file)

# Add a worksheet to my_book, named "data_summary"
createSheet(my_book, "data_summary")

# Create data frame: summ
sheets <- getSheets(my_book)[1:3]
dims <- sapply(sheets, function(x) dim(readWorksheet(my_book, sheet = x)), USE.NAMES = FALSE)
summ <- data.frame(sheets = sheets,
                   nrows = dims[1, ],
                   ncols = dims[2, ])

# Add data in summ to "data_summary" sheet
writeWorksheet(my_book, summ, sheet = "data_summary")

# Save workbook as summary.xlsx
saveWorkbook(my_book, out_xlsx_file)
```

```{r}
# Build connection to urbanpop.xlsx: my_book
urbanpop_summary <- file.path("..", "00_Datasets", "urbanpop_summary.xlsx")
urbanpop_renamed <- file.path("..", "00_Datasets", "renamed.xlsx")


my_book <- loadWorkbook(urbanpop_summary)

# Rename "data_summary" sheet to "summary"
renameSheet(my_book, "data_summary", "summary")

# Print out sheets of my_book
getSheets(my_book)

# Save workbook to "renamed.xlsx"
saveWorkbook(my_book, file = urbanpop_renamed)
```

### **`07-Removing sheets`**

-   Load the `XLConnect` package.

-   Build a connection to `"renamed.xlsx"`, the Excel file that you've built in the previous exercise; it's available in your working directory. Store this connection as `my_book`.

-   Use `removeSheet()` to remove the fourth sheet from `my_book`. The sheet name is `"summary"`. Recall that `removeSheet()` accepts either the index or the name of the sheet as the second argument.

-   Save the resulting workbook, `my_book`, to a file `"clean.xlsx"`.

```{r}
# output path
urbanpop_cleaned <- file.path("..", "00_Datasets", "clean.xlsx")

# Load the XLConnect package
library(XLConnect)

# Build connection to renamed.xlsx: my_book
my_book <- loadWorkbook(urbanpop_renamed)

# Remove the fourth sheet
removeSheet(my_book, sheet = "summary")

# Save workbook to "clean.xlsx"
saveWorkbook(my_book, urbanpop_cleaned)
```

### `The End`

 



 



 



 



 



 
