---
title: "Data cleaning and summarizing with dplyr"
author: "Mohamad Osman"
date: '2022-07-14'
output: rmarkdown::github_document
---

# **`Section 01: Data cleaning and summarizing with dplyr`**

### **`01-Filtering rows`**

The `vote` column in the dataset has a number that represents that country's vote:

-   **1** = Yes

-   **2** = Abstain

-   **3** = No

-   **8** = Not present

-   **9** = Not a member

One step of data cleaning is removing observations (rows) that you're not interested in. In this case, you want to remove "Not present" and "Not a member".

```{r}
library(dplyr)
library(tidyverse)
```

```{r}
votes_file <- file.path("..", "00_Datasets", "votes.rds")
votes <- readRDS(votes_file)
```

-   Load the `dplyr` package.

-   Print the `votes` table.

-   Filter out rows where the vote recorded is "not present" or "not a member", leaving cases where it is "yes", "abstain", or "no".

```{r}
    # Load the dplyr package
library(dplyr)

# Print the votes dataset
head(votes)

# Filter for votes that are "yes", "abstain", or "no"
votes %>%
    filter(vote <= 3) %>%
    head()
```

### **`02-Adding a year column`**

The next step of data cleaning is manipulating your variables (columns) to make them more informative.

In this case, you have a `session` column that is hard to interpret intuitively. But since the UN started voting in 1946, and holds one session per year, you can get the year of a UN resolution by adding 1945 to the `session` number.

```{r}
# Add another %>% step to add a year column
votes %>%
  filter(vote <= 3) %>%
  mutate(year = session + 1945) %>%
  head()
```

### **`03-Adding a country column`**

The country codes in the `ccode` column are what's called [**Correlates of War codes**](https://cow.dss.ucdavis.edu/data-sets/cow-country-codes). This isn't ideal for an analysis, since you'd like to work with recognizable country names.

You can use the `countrycode` package to translate. For example:

    library(countrycode)

    # Translate the country code 2
    > countrycode(2, "cown", "country.name")
    [1] "United States"

    # Translate multiple country codes
    > countrycode(c(2, 20, 40), "cown", "country.name")
    [1] "United States" "Canada"        "Cuba"

-   Load the `countrycode` package.

-   Convert the country code 100 to its country name.

-   Add a new `country` column in your `mutate()` statement containing country names, using the `countrycode()` function to translate from the `ccode` column. Save the result to `votes_processed`.

```{r}
# Load the countrycode package
library(countrycode)

# Convert country code 100
countrycode(100, "cown", "country.name")

# Add a country column within the mutate: votes_processed
votes_processed <- votes %>%
  filter(vote <= 3) %>%
  mutate(year = session + 1945, 
         country = countrycode(ccode, "cown", "country.name")) 

  
head(votes_processed)
```

### 
**`04-Summarizing the full dataset`**

In this analysis, you're going to focus on "% of votes that are yes" as a metric for the "agreeableness" of countries.

You'll start by finding this summary for the entire dataset: the fraction of all votes in their history that were "yes". Note that within your call to `summarize()`, you can use `n()` to find the total number of votes and `mean(vote == 1)` to find the fraction of "yes" votes.

-   Print the `votes_processed` dataset that you created in the previous exercise.

-   Summarize the dataset using the `summarize()` function to create two columns:

    -   `total`: with the number of votes

    -   `percent_yes`: the percentage of "yes" votes

```{r}
# Print votes_processed
head(votes_processed)

# Find total and fraction of "yes" votes
votes_processed %>%
    summarise(total = n(),
              percent_yes = mean(vote == 1))
```

### `05-Summarizing by year`

The `summarize()` function is especially useful because it can be used within *groups*.

For example, you might like to know how much the average "agreeableness" of countries changed from year to year. To examine this, you can use `group_by()` to perform your summary not for the entire dataset, but within each year.

-   Add a `group_by()` to your code to `summarize()` within each year.

```{r}
# Change this code to summarize by year
votes_processed %>%
  group_by(year) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))
```

### **`06-Summarizing by country`**

In the last exercise, you performed a summary of the votes within each year. You could instead `summarize()` within each country, which would let you compare voting patterns between countries.

```{r}
# Summarize by country: by_country
by_country <- votes_processed %>%
  group_by(country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

head(by_country)
```

### **`07-Sorting by percentage of "yes" votes`**

Now that you've summarized the dataset by country, you can start examining it and answering interesting questions.

For example, you might be especially interested in the countries that voted "yes" least often, or the ones that voted "yes" most often.

-   Print the `by_country` dataset created in the last exercise.

-   Use `arrange()` to sort the countries in ascending order of `percent_yes`.

-   Arrange the countries by the same variable, but in descending order.

```{r}
# You have the votes summarized by country
by_country <- votes_processed %>%
  group_by(country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# Print the by_country dataset
by_country

# Sort in ascending order of percent_yes
by_country %>% 
  arrange(percent_yes)

# Now sort in descending order
by_country %>% 
  arrange(desc(percent_yes))
```

### **`08-Filtering summarized output`**

In the last exercise, you may have noticed that the country that voted least frequently, Zanzibar, had only 2 votes in the entire dataset. You certainly can't make any substantial conclusions based on that data!

Typically in a progressive analysis, when you find that a few of your observations have very little data while others have plenty, you set some threshold to filter them out

-   Use `filter()` to remove from the sorted data countries that have fewer than 100 votes.

```{r}
# Filter out countries with fewer than 100 votes
by_country %>%
  arrange(percent_yes) %>%
  filter(total >= 100)

```

### `THE END`









