---
title: "Joining and tidying"
author: "Mohamad Osman"
date: '2022-07-15'
output: rmarkdown::github_document
---

# Section 04: **Joining and tidying**

```{r}
library(countrycode)
library(dplyr)
```

```{r}
description_file <- file.path("..", "00_Datasets", "descriptions.rds")
descriptions <- readRDS(description_file)

votes_file <- file.path("..", "00_Datasets", "votes.rds")
votes <- readRDS(votes_file)

# Add a country column within the mutate: votes_processed
votes_processed <- votes %>%
  filter(vote <= 3) %>%
  mutate(year = session + 1945, 
         country = countrycode(ccode, "cown", "country.name"))
```

### **`01-Joining datasets with inner_join`**

In the first chapter, you created the `votes_processed` dataset, containing information about each country's votes. You'll now combine that with the new `descriptions` dataset, which includes topic information about each country, so that you can analyze votes within particular topics.

To do this, you'll make use of the `inner_join()` function from `dplyr`.

```{r}
# Load dplyr package
#library(dplyr)

# Print the votes_processed dataset
votes_processed

# Print the descriptions dataset
descriptions

# Join them together based on the "rcid" and "session" columns
votes_joined <- votes_processed %>%
    inner_join(descriptions, by = c("rcid", "session"))

```

### **`02-Filtering the joined dataset`**

There are six columns in the `descriptions` dataset (and therefore in the new joined dataset) that describe the topic of a resolution:

1.  **me**: Palestinian conflict

2.  **nu**: Nuclear weapons and nuclear material

3.  **di**: Arms control and disarmament

4.  **hr**: Human rights

5.  **co**: Colonialism

6.  **ec**: Economic development

Each contains a `1` if the resolution is related to this topic and a `0` otherwise.

-   Filter the `votes_joined` dataset for votes relating to colonialism.

```{r}
# Filter for votes related to colonialism
votes_joined %>%
    filter(co >= 1)
```

### **`03-Visualizing colonialism votes`**

In an earlier exercise, you graphed the percentage of votes each year where the US voted "yes". Now you'll create that same graph, but only for votes related to colonialism.

-   Load the `ggplot2` package.

-   Filter the `votes_joined` dataset for only votes by the United States relating to colonialism, then `summarize()` the percentage of votes that are "yes" within each `year`. Name the resulting column `percent_yes` and save the entire data frame as `US_co_by_year`.

-   Add a `geom_line()` layer to your `ggplot()` call to create a line graph of the percentage of "yes" votes on colonialism (`percent_yes`) cast by the US over time.

```{r}
# Load the ggplot2 package
library(ggplot2)

# Filter, then summarize by year: US_co_by_year
US_co_by_year <- votes_joined %>%
    filter(country == "United States",
           co > 0) %>%
    group_by(year) %>%
    summarize(percent_yes = mean(vote == 1))


    
# Graph the % of "yes" votes over time
ggplot(US_co_by_year, aes(year, percent_yes)) +
  geom_line()
```

### **`04-Using gather to tidy a dataset`**

In order to represent the joined vote-topic data in a tidy form so we can analyze and graph by topic, we need to transform the data so that each row has one combination of country-vote-topic. This will change the data from having six columns (`me`, `nu`, `di`, `hr`, `co`, `ec`) to having two columns (`topic` and `has_topic`).

-   Load the `tidyr` package.

-   Gather the six topic columns in `votes_joined` into one column called `topic` (containing one of `me`, `nu`, etc.) and a column called `has_topic` (containing 0 or 1). Print the result without saving it.

-   You don't actually care about the cases where `has_topic` is 0. Perform the `gather()` operation again, but this time also filter for only the rows where the topic in `topic` describes the vote. Save the result as `votes_gathered`.

```{r}
# Load the tidyr package
library(tidyr)

# Gather the six me/nu/di/hr/co/ec columns
votes_joined %>%
    gather(topic, has_topic, me:ec) %>%
    head()


# Perform gather again, then filter
votes_gathered <- votes_joined %>%
    gather(topic, has_topic, me:ec) %>%
    filter(has_topic == 1)

head(votes_gathered)
```

### **`05-Recoding the topics`**

There's one more step of data cleaning to make this more interpretable. Right now, topics are represented by two-letter codes:

1.  **me**: Palestinian conflict

2.  **nu**: Nuclear weapons and nuclear material

3.  **di**: Arms control and disarmament

4.  **hr**: Human rights

5.  **co**: Colonialism

6.  **ec**: Economic development

So that you can interpret the data more easily, recode the data to replace these codes with their full name. You can do that with `dplyr`'s `recode()` function, which replaces values with ones you specify:

    example <- c("apple", "banana", "apple", "orange")
    recode(example,
           apple = "plum",
           banana = "grape"

-   Use the `recode()` function from `dplyr` in a `mutate()` to replace each two-letter code in the `votes_gathered` data frame with the corresponding full name. Save this as `votes_tidied`.

```{r}
# Replace the two-letter codes in topic: votes_tidied
votes_tidied <- votes_gathered %>%
  mutate(topic = recode(topic,
                        me = "Palestinian conflict",
                        nu = "Nuclear weapons and nuclear material",
                        di = "Arms control and disarmament",
                        hr = "Human rights",
                        co = "Colonialism",
                        ec = "Economic development"))

```

### **`06-Summarize by country, year, and topic`**

In previous exercises, you summarized the votes dataset by country, by year, and by country-year combination.

Now that you have `topic` as an additional variable, you can summarize the votes for each combination of country, year, *and* topic (e.g. for the United States in 2013 on the topic of nuclear weapons.)

-   Print the `votes_tidied` dataset to the console.

-   In a single `summarize()` call, compute both the total number of votes (`total`) and the percentage of "yes" votes (`percent_yes`) for each combination of country, year, and topic. Save this as `by_country_year_topic`. Make sure that you `ungroup()` after summarizing.

-   Print the `by_country_year_topic` dataset to the console.

```{r}
# Print votes_tidied
votes_tidied

# Summarize the percentage "yes" per country-year-topic
by_country_year_topic <- votes_tidied %>%
        group_by(country, year, topic)  %>%
        summarise(total = n(),
                  percent_yes = mean(vote == 1)) %>%
        ungroup()

# Print by_country_year_topic
by_country_year_topic
```

### **`07-Visualizing trends in topics for one country`**

You can now visualize the trends in percentage "yes" over time for all six topics side-by-side. Here, you'll visualize them just for the United States.

-   Load the `ggplot2` package.

-   Filter the `by_country_year_topic` dataset for just the United States and save the result as `US_by_country_year_topic`.

-   Use this dataset to graph the percentage "yes" over time for the United States, faceting by topic.

```{r}
# Load the ggplot2 package
library(ggplot2)

# Filter by_country_year_topic for just the US
US_by_country_year_topic <- by_country_year_topic %>%
    filter(country == "United States")

# Plot % yes over time for the US, faceting by topic
ggplot(US_by_country_year_topic, aes(year, percent_yes)) +
    geom_line() +
    facet_wrap(~ topic)
```

### **`08-Nesting by topic and country`**

In the last chapter, you constructed a linear model for each country by nesting the data in each country, fitting a model to each dataset, then tidying each model with `broom` and unnesting the coefficients. The code looked something like this:

    country_coefficients <- by_year_country %>%
      nest(-country) %>%
      mutate(model = map(data, ~ lm(percent_yes ~ year, data = .)),
             tidied = map(model, tidy)) %>%
      unnest(tidied)

Now, you'll again be modeling change in "percentage" yes over time, but instead of fitting one model for each country, you'll fit one for each combination of country and topic.

-   Load the `purrr`, `tidyr`, and `broom` packages.

-   Print the `by_country_year_topic` dataset to the console.

-   Fit a linear model within each country and topic in this dataset, saving the result as `country_topic_coefficients`. You can use the provided code as a starting point.

-   Print the `country_topic_coefficients` dataset to the console.

```{r}
# Load purrr, tidyr, and broom
library(purrr)
library(tidyr)
library(broom)

# Print by_country_year_topic
by_country_year_topic

# Fit model on the by_country_year_topic dataset
country_topic_coefficients <- by_country_year_topic %>%
        nest(-country, -topic) %>%
        mutate(model = map(data,  ~ lm(percent_yes ~ year, data = .)),
               tidied = map(model, tidy)) %>%
        unnest(tidied)

# Print country_topic_coefficients
country_topic_coefficients
```

### `09-Interpreting tidy models`

Now you have both the slope and intercept terms for each model. Just as you did in the last chapter with the tidied coefficients, you'll need to filter for only the slope terms.

You'll also have to extract only cases that are statistically significant, which means adjusting the p-value for the number of models, and then filtering to include only significant changes.

-   Filter the `country_topic_coefficients` data to include only the slope term.

-   Add a `p.adjusted` column containing adjusted p-values (using the `p.adjust()` function).

-   Filter for only adjusted p-values less than .05.

-   Save the result as `country_topic_filtered`.

```{r}
# Create country_topic_filtered
country_topic_filtered <- country_topic_coefficients %>%
    filter(term == "year") %>%
    mutate(p.adjusted = p.adjust(p.value)) %>%
    filter(p.adjusted < 0.05)
```

### 
**`10-Checking models visually`**

In the last exercise, you found that over its history, Vanuatu (an island nation in the Pacific Ocean) sharply changed its pattern of voting on the topic of Palestinian conflict.

Let's examine this country's voting patterns more closely. Recall that the `by_country_year_topic` dataset contained one row for each combination of country, year, and topic. You can use that to create a plot of Vanuatu's voting, faceted by topic.

-   Filter the `by_country_year_topic` variable for only Vanuatu's votes to create a `vanuatu_by_country_year_topic` object.

-   Create a plot with `year` on the x-axis and `percent_yes` on the y-axis, and facet by topic.

```{r}
# Create vanuatu_by_country_year_topic
vanuatu_by_country_year_topic <- by_country_year_topic %>%
  filter(country == "Vanuatu")

# Plot of percentage "yes" over time, faceted by topic
ggplot(vanuatu_by_country_year_topic, aes(year, percent_yes)) +
  geom_line() +
  facet_wrap(~ topic)
```

### `The End`




























