---
title: "Cleaning Data in R"
author: "Mohamad Osman"
date: '2022-06-26'
output: rmarkdown::github_document
---

# Section 01: **Common Data Problems**

### **`01-Converting data types`**

-   Examine the data types of the columns of `bike_share_rides`.

-   Get a summary of the `user_birth_year` column of `bike_share_rides`.

```{r}
library(tidyverse)
library(dplyr)
library(assertive)
library(stringr)
```

```{r}

# file path
bike_share_file <- file.path("..", "00_Datasets", "bike_share_rides_ch1_1.rds") 

# reading rds file
bike_share_rides <- readRDS(bike_share_file)


# Glimpse at bike_share_rides
glimpse(bike_share_rides)

# Summary of user_birth_year
summary(bike_share_rides$user_birth_year)
```

-   Add a new column to `bike_share_rides` called `user_birth_year_fct`, which contains `user_birth_year`, converted to a factor.

-   Assert that the `user_birth_year_fct` is a factor to confirm the conversion.

```{r}
# Glimpse at bike_share_rides
glimpse(bike_share_rides)

# Summary of user_birth_year
#summary(bike_share_rides$user_birth_year)

# Convert user_birth_year to factor: user_birth_year_fct
bike_share_rides <- bike_share_rides %>%
  mutate(user_birth_year_fct = as.factor(user_birth_year))

# Assert user_birth_year_fct is a factor
assert_is_factor(bike_share_rides$user_birth_year_fct)

# Summary of user_birth_year_fct
summary(bike_share_rides$user_birth_year_fct)
```

### **`02-Trimming strings`**

-   Use `str_remove()` to remove `"minutes"` from the `duration` column of `bike_share_rides`. Add this as a new column called `duration_trimmed`.

-   Convert the `duration_trimmed` column to a numeric type and add this as a new column called `duration_mins`.

-   Glimpse at `bike_share_rides` and assert that the `duration_mins` column is `numeric`.

-   Calculate the mean of `duration_mins`.

```{r}
bike_share_rides <- bike_share_rides %>%
  # Remove 'minutes' from duration: duration_trimmed
  mutate(duration_trimmed = str_remove(duration, "minutes"),
         # Convert duration_trimmed to numeric: duration_mins
         duration_mins = as.numeric(duration_trimmed))

# Glimpse at bike_share_rides
glimpse(bike_share_rides)

# Assert duration_mins is numeric
assert_is_numeric(bike_share_rides$duration_mins)

# Calculate mean duration
mean(bike_share_rides$duration_mins)
```

### **`03-Ride duration constraints`**

1.  Create a three-bin histogram of the `duration_min` column of `bike_share_rides` using `ggplot2` to identify if there is out-of-range data.

```{r}
bike_share_file <- file.path("..", "00_Datasets", "bike_share_rides_nn.csv") 
bike_share_rides <- read.csv(bike_share_file)
```

```{r}
# Create breaks
breaks <- c(min(bike_share_rides$duration_min), 0, 1440,
            max(bike_share_rides$duration_min))

# Create a histogram of duration_min
ggplot(bike_share_rides, aes(duration_min)) +
  geom_histogram(breaks = breaks)
```

-   Replace the values of `duration_min` that are greater than `1440` minutes (24 hours) with `1440`. Add this to `bike_share_rides` as a **new** column called `duration_min_const`.

-   Assert that all values of `duration_min_const` are between `0` and `1440`.

```{r}
# Create breaks
breaks <- c(min(bike_share_rides$duration_min), 0, 1440, max(bike_share_rides$duration_min))

# Create a histogram of duration_min
ggplot(bike_share_rides, aes(duration_min)) +
  geom_histogram(breaks = breaks)

# duration_min_const: replace vals of duration_min > 1440 with 1440
bike_share_rides <- bike_share_rides %>%
  mutate(duration_min_const = replace(duration_min, duration_min > 1440, 1440))

# Make sure all values of duration_min_const are between 0 and 1440
assert_all_are_in_closed_range(bike_share_rides$duration_min_const, lower = 0, upper = 1440)

# Create a histogram of duration_min
ggplot(bike_share_rides, aes(duration_min_const)) +
  geom_histogram(breaks = breaks)
```

### **`04-Back to the future`**

-   Convert the `date` column of `bike_share_rides` from `character` to the `Date` data type.

-   Assert that all values in the `date` column happened sometime in the past and not in the future.

-   Filter `bike_share_rides` to get only the rides from the past or today, and save this as `bike_share_rides_past`.

-   Assert that the `dates` in `bike_share_rides_past` occurred only in the past.

```{r}
bike_share_file <- file.path("..", "00_Datasets", "bike_share_rides_date_version.csv") 
bike_share_rides <- read.csv(bike_share_file)
bike_share_rides$date <- as.Date(bike_share_rides$date, format = "%m/%d/%Y")
str(bike_share_rides$date[1:10])
```

```{r}
library(lubridate)

# Convert date to Date type
bike_share_rides <- bike_share_rides %>%
  mutate(date = as.Date(date))


# Make sure all dates are in the past
#assert_all_are_in_past(bike_share_rides$date)


# Filter for rides that occurred before or on today's date
bike_share_rides_past <- bike_share_rides %>%
  filter(date <= today())

# Make sure all dates from bike_share_rides_past are in the past
assert_all_are_in_past(bike_share_rides_past$date)

```

### **`05-Full duplicates`**

-   Get the total number of full duplicates in `bike_share_rides`.

-   Remove all full duplicates from `bike_share_rides` and save the new data frame as `bike_share_rides_unique.`

-   Get the total number of full duplicates in the new `bike_share_rides_unique` data frame.

```{r}
bike_share_file <- file.path("..", "00_Datasets", "bike_share_rides_full_duplicated.csv") 

bike_share_rides <- read.csv(bike_share_file)



# Count the number of full duplicates
sum(duplicated(bike_share_rides))

# Remove duplicates
bike_share_rides_unique <- distinct(bike_share_rides)


# Count the full duplicates in bike_share_rides_unique
sum(duplicated(bike_share_rides_unique))
```

### **`06-Removing partial duplicates`**

-   Count the number of occurrences of each `ride_id`.

-   Filter for `ride_id`s that occur multiple times.

```{r}
bike_share_file <- file.path("..", "00_Datasets", "bike_share_rides_partial_duplicated.csv") 

bike_share_rides <- read.csv(bike_share_file)


# Find duplicated ride_ids
bike_share_rides %>% 
  # Count the number of occurrences of each ride_id
  count(ride_id) %>% 
  # Filter for rows with a count > 1
  filter(n > 1)
```

-   Remove full and partial duplicates from `bike_share_rides` based on `ride_id` only, keeping all columns.

-   Store this as `bike_share_rides_unique`.

-   Find the duplicated `ride_id`s in `bike_share_rides_unique`.

```{r}
# Find duplicated ride_ids
bike_share_rides %>% 
  count(ride_id) %>% 
  filter(n > 1)

# Remove full and partial duplicates
bike_share_rides_unique <- bike_share_rides %>%
  # Only based on ride_id instead of all cols
  distinct(ride_id, .keep_all = TRUE)

# Find duplicated ride_ids in bike_share_rides_unique
bike_share_rides_unique %>%
  # Count the number of occurrences of each ride_id
  count(ride_id) %>%
  # Filter for rows with a count > 1
  filter(n > 1)
```

### **07-Aggregating partial duplicates**

-   Group `bike_share_rides` by `ride_id` and `date`.

-   Add a column called `duration_min_avg` that contains the mean ride duration for the row's `ride_id` and `date`.

-   Remove duplicates based on `ride_id` and `date`, keeping all columns of the data frame.

-   Remove the `duration_min` column.

```{r}
bike_share_rides %>%
  # Group by ride_id and date
  group_by(ride_id, date) %>%
  # Add duration_min_avg column
  mutate(duration_min_avg = mean(duration_min) ) %>%
  # Remove duplicates based on ride_id and date, keep all cols
  distinct(ride_id, date, .keep_all = TRUE) %>%
  # Remove duration_min column
  select(-duration_min)
```

### `The End`
