---
title: "Categorical and Text Data"
author: "Mohamad Osman"
date: '2022-06-26'
output: rmarkdown::github_document
---

# Section 02: **Categorical and Text Data**

### **`01-Not a member (membership constraints)`**

-   Count the number of occurrences of each `dest_size` in `sfo_survey`.

```{r}
library(tidyverse)
library(dplyr)
```

```{r}
# file path
sfo_survey_file <- file.path("..", "00_Datasets", "sfo_survey_ch2_1.rds") 

# reading rds file
sfo_survey <- readRDS(sfo_survey_file)

# Count the number of occurrences of dest_size
sfo_survey %>%
  count(dest_size)
```

-   Use the correct type of filtering join on the `sfo_survey` data frame and the `dest_sizes` data frame to get the rows of `sfo_survey` with invalid `dest_size` values.

-   Get the `id`, `airline`, `destination`, and `dest_size` columns.

```{r}
# Create a truth table for validation 
dest_size <- c("Small", "Medium", "Large", "Hub")
passengers_per_day <- c("0-20K", "20K-70K", "70K-100K", "100K+")

dest_sizes <- data.frame(dest_size, passengers_per_day, stringsAsFactors = FALSE)

```

```{r}
# Find bad dest_size rows
sfo_survey %>% 
  # Join with dest_sizes data frame to get bad dest_size rows
  anti_join(dest_sizes) %>%
  # Select id, airline, destination, and dest_size cols
  select(id, airline, destination, dest_size)
```

-   Use the correct filtering join on `sfo_survey` and `dest_sizes` to get the rows of `sfo_survey` that have a valid `dest_size`.

-   Count the number of times that each `dest_size` occurs to make sure there are no invalid values left behind

```{r}
# Remove bad dest_size rows
sfo_survey %>% 
  # Join with dest_sizes
  semi_join(dest_sizes) %>%
  # Count the number of each dest_size
  count(dest_size)
```

### **`02-Identifying inconsistency`**

In this exercise, you'll continue working with the `sfo_survey` dataset. You'll examine the `dest_size` column again as well as the `cleanliness` column and determine what kind of issues, if any, these two categorical variables face.

`dplyr` and is loaded and `sfo_survey` is available.

-   Count the number of occurrences of each category of the `dest_size` variable of `sfo_survey`.

```{r}
# Count dest_size
sfo_survey %>%
  count(dest_size)
```

The categories in `dest_size` have inconsistent white space.

-   Count the number of occurrences of each category of the `cleanliness` variable of `sfo_survey`.

```{r}
# Count dest_size
sfo_survey %>%
  count(dest_size)

# Count cleanliness
sfo_survey %>%
  count(cleanliness)
```

### **`03-Correcting inconsistency`**

-   Add a column to `sfo_survey` called `dest_size_trimmed` that contains the values in the `dest_size` column with all leading and trailing whitespace removed.

-   Add another column called `cleanliness_lower` that contains the values in the `cleanliness` column converted to all lowercase.

-   Count the number of occurrences of each category in `dest_size_trimmed`.

-   Count the number of occurrences of each category in `cleanliness_lower`.

```{r}
# Add new columns to sfo_survey
sfo_survey <- sfo_survey %>%
  # dest_size_trimmed: dest_size without whitespace
  mutate(dest_size_trimmed = str_trim(dest_size),
         # cleanliness_lower: cleanliness converted to lowercase
         cleanliness_lower = str_to_lower(cleanliness))

# Count values of dest_size_trimmed
sfo_survey %>%
  count(dest_size_trimmed)

# Count values of cleanliness_lower
sfo_survey %>%
  count(cleanliness_lower)
```

### **`04-Collapsing categories`**

-   Count the categories of `dest_region`.

```{r}
# Count categories of dest_region
sfo_survey %>%
  count(dest_region)
```

`"EU"`, `"eur"`, and `"Europ"` need to be collapsed to `"Europe"`.

-   Create a vector called `europe_categories` containing the three values of `dest_region` that need to be collapsed.

-   Add a new column to `sfo_survey` called `dest_region_collapsed` that contains the values from the `dest_region` column, except the categories stored in `europe_categories` should be collapsed to `Europe`.

-   Count the categories of `dest_region_collapsed`.

```{r}
library(forcats)

# Count categories of dest_region
sfo_survey %>%
  count(dest_region)

# Categories to map to Europe
europe_categories <- c("EU", "eur", "Europ")

# Add a new col dest_region_collapsed
sfo_survey %>%
  # Map all categories in europe_categories to Europe
  mutate(dest_region_collapsed = fct_collapse(dest_region, 
                                     Europe = europe_categories)) %>%
  # Count categories of dest_region_collapsed
  count(dest_region_collapsed)
```

### **`05-Detecting inconsistent text data`**

`dplyr` and `stringr` are loaded, and `sfo_survey` is available.

-   Filter for rows with phone numbers that contain `"-"`s.

-   Filter for rows with phone numbers that contain `"("`, or `")"`. Remember to use `fixed()` when searching for parentheses.

```{r}
library(stringr)

# file path
sfo_survey_file <- file.path("..", "00_Datasets", "sfo_survey_phone.csv") 

# reading rds file
sfo_survey <- read.csv(sfo_survey_file)
```

```{r}
# Filter for rows with "-" in the phone column
sfo_survey %>%
  filter(str_detect(phone, "-"))
```

-   Filter for rows with phone numbers that contain `"("`, or `")"`. Remember to use `fixed()` when searching for parentheses.

```{r}
# Filter for rows with "(" or ")" in the phone column
sfo_survey %>%
  filter(str_detect(phone, fixed("(")) | str_detect(phone, fixed(")")))

```

### **`06-Replacing and removing`**

-   Remove opening and closing parentheses from the `phone` column. Store this as a variable called `phone_no_parens`. Remember to use `fixed()`!

-   Add a new column to `sfo_survey` called `phone_no_parens` that contains the contents of `phone_no_parens`.

-   Create a new column of `sfo_survey` called `phone_clean` containing the values of `phone_no_parens` with all hyphens replaced with spaces.

```{r}
# Remove parentheses from phone column
phone_no_parens <- sfo_survey$phone %>%
  # Remove "("s
  str_remove_all(fixed("(")) %>%
  # Remove ")"s
  str_remove_all(fixed(")"))

# Add phone_no_parens as column
sfo_survey %>%
  mutate(phone_no_parens = phone_no_parens,
  # Replace all hyphens in phone_no_parens with spaces
         phone_clean = str_replace_all(phone_no_parens, "-", " "))
```

### **`07-Invalid phone numbers`**

-   Examine the invalid `phone` numbers by filtering for numbers whose length is not equal to 12.

-   Remove the rows with invalid numbers by filtering for numbers with a length of exactly 12.

```{r}
# Check out the invalid numbers
sfo_survey %>%
  filter(str_length(phone) != 12)

# Remove rows with invalid numbers
sfo_survey %>%
  filter(str_length(phone) == 12)

```

### `The End`



 

 

 








 
