---
title: "Advanced Data Problems"
author: "Mohamad Osman"
date: '2022-06-26'
output: rmarkdown::github_document
---

# `Section 03: Advanced Data Problems`

### `01-Date uniformity`

-   Take a look at the head of `accounts` to get a sense of the data you're working with.

```{r}
# file path 
accounts_file = file.path("..", "00_Datasets", "ch3_1_accounts.rds")

accounts <- readRDS(accounts_file)
```

-   Take a look at the head of `accounts` to get a sense of the data you're working with.

```{r}
library(lubridate)
library(dplyr)
library(ggplot2)
```

```{r}
# Check out the accounts data frame
head(accounts)
```

```{r}
# Check out the accounts data frame
head(accounts)

# Define the date formats
formats <- c("%Y-%m-%d", "%B %d, %Y")

# Convert dates to the same format
accounts %>%
  mutate(date_opened_clean = parse_date_time(accounts$date_opened, formats))
```

-   Create a scatter plot with `date_opened` on the x-axis and `total` on the y-axis.

```{r}
accounts <- accounts %>%
  mutate(date_opened = parse_date_time(date_opened, formats))
```

```{r}
# Scatter plot of opening date and total amount
accounts %>%
  ggplot(aes(x = date_opened, y = total)) +
  geom_point()
```

```{r}
accounts_file_path <- file.path("..", "00_Datasets", "account_offices.csv")
account_offices <- read.csv(accounts_file_path)
```

```{r}
# Scatter plot of opening date and total amount
accounts %>%
  ggplot(aes(x = date_opened, y = total)) +
  geom_point()

# Left join accounts and account_offices by id
accounts %>%
  left_join(account_offices, by = c("id"))
```

### **`02-Currency uniformity`**

The formula to convert yen to dollars is `USD = JPY / 104`.

`dplyr` and `ggplot2` are loaded and the `accounts` and `account_offices` data frames are available.

-   Convert the `total`s from the Tokyo office from yen to dollars, and keep the `total` from the New York office in dollars. Store this as a new column called `total_usd`.

-   Create a scatter plot of your new uniform data using `date_opened` on the x-axis and `total_usd` on the y-axis.

```{r}
# Scatter plot of opening date and total amount
accounts %>%
  ggplot(aes(x = date_opened, y = total)) +
  geom_point()

# Left join accounts to account_offices by id
accounts %>%
  left_join(account_offices, by = "id") %>%
  # Convert totals from the Tokyo office to USD
  mutate(total_usd = ifelse(office == "Tokyo", total/104, total)) %>%
  # Scatter plot of opening date vs total_usd
  ggplot(aes(x = date_opened, y = total_usd)) +
  geom_point()
```

### 
**`03-Validating totals`**

-   Create a new column called `theoretical_total` that contains the sum of the amounts in each fund.

-   Find the accounts where the `total` doesn't match the `theoretical_total`

```{r}
file_path = file.path("..", "00_Datasets", "accounts_with_funds.csv")
accounts <- read.csv(file_path)
```

```{r}
# Find invalid totals
accounts %>%
  # theoretical_total: sum of the three funds
  mutate(theoretical_total = fund_A + fund_B + fund_C) %>%
  # Find accounts where total doesn't match theoretical_total
  filter(total != theoretical_total)
```

### **`04-Validating age`**

-   Create a new column called `theoretical_age` that contains the age of each account based on the `date_opened`.

-   Find the accounts where the `acct_age` doesn't match the `theoretical_age`.

```{r}
# read the csv file.
file_path = file.path("..", "00_Datasets", "accounts_with_ages.csv")
accounts <- read.csv(file_path)
accounts
# change the format of the date opended. 
accounts <- accounts %>%
  mutate(date_opened = parse_date_time(date_opened, "%m/%d/$Y"))

# Find invalid acct_age
accounts %>%
  # theoretical_age: age of acct based on date_opened
  mutate(theoretical_age = floor(as.numeric(date_opened %--% today(), "years"))) %>%
  # Filter for rows where acct_age is different from theoretical_age
  filter(theoretical_age != acct_age)
```

### **`05-Visualizing missing data`**

-   Visualize the missing values in `accounts` by column using a function from the `visdat` package.

```{r}
library(visdat)
```

```{r}
# read account 
account_m_file <- file.path("..", "00_Datasets", "accounts_missing_values.csv")
accounts <- read.csv(account_m_file)

# Visualize the missing values by column
vis_miss(accounts)
```

-   Add a logical column to `accounts` called `missing_inv` that indicates whether each row is missing the `inv_amount` or not.

-   Group by `missing_inv`.

-   Calculate the mean age for each group of `missing_inv`.

```{r}
# Visualize the missing values by column
vis_miss(accounts)

accounts %>%
  # missing_inv: Is inv_amount missing?
  mutate(missing_inv = is.na(inv_amount)) %>%
  # Group by missing_inv
  group_by(missing_inv) %>%
  # Calculate mean age for each missing_inv group
  summarize(avg_age = mean(age))
```

-   Sort `accounts` by `age`.

-   Visualize missing data by column.

```{r}
# Visualize the missing values by column
vis_miss(accounts)

accounts %>%
  # missing_inv: Is inv_amount missing?
  mutate(missing_inv = is.na(inv_amount)) %>%
  # Group by missing_inv
  group_by(missing_inv) %>%
  # Calculate mean age for each missing_inv group
  summarize(avg_age = mean(age))

# Sort by age and visualize missing vals
#accounts %>%
  #arrange(age)%>%
  #vis_miss(is.na(accounts)))
```

### **06-Treating missing data**

In this exercise, you're working with another version of the `accounts` data that contains missing values for both the `cust_id` and `acct_amount` columns.

You want to figure out how many unique customers the bank has, as well as the average amount held by customers. You know that rows with missing `cust_id` don't really help you, and that on average, the `acct_amount` is usually 5 times the amount of `inv_amount`.

In this exercise, you will drop rows of `accounts` with missing `cust_id`s, and impute missing values of `inv_amount` with some domain knowledge. `dplyr` and `assertive` are loaded and `accounts` is available.

Filter `accounts` to remove rows with missing `cust_id`s and save as `accounts_clean`.

```{r}
# read account 
account_m_file <- file.path("..", "00_Datasets", "accounts_missing_values.csv")
accounts <- read.csv(account_m_file)
```

```{r}
# Create accounts_clean
accounts_clean <- accounts %>%
  # Filter to remove rows with missing cust_id
  filter(!is.na(cust_id))

accounts_clean
```

-   Create a new column called `acct_amount_filled`, which contains the values of `acct_amount`, except all `NA` values should be replaced with 5 times the amount in `inv_amount`.

```{r}
# Create accounts_clean
accounts_clean <- accounts %>%
  # Filter to remove rows with missing cust_id
  filter(!is.na(cust_id)) %>%
  # Add new col acct_amount_filled with replaced NAs
  mutate(acct_amount_filled = ifelse(is.na(acct_amount), 5 * inv_amount, acct_amount))

accounts_clean
```

-   Assert that there are no missing values in the `acct_amount_filled` column of `accounts_clean`.

-   Assert that there are no missing values in the `cust_id` column of `accounts_clean`.

```{r}
library(assertive)

# Create accounts_clean
accounts_clean <- accounts %>%
  # Filter to remove rows with missing cust_id
  filter(!is.na(cust_id)) %>%
  # Add new col acct_amount_filled with replaced NAs
  mutate(acct_amount_filled = ifelse(is.na(acct_amount), inv_amount * 5, acct_amount))

# Assert that cust_id has no missing vals
assert_all_are_not_na(accounts_clean$cust_id)

# Assert that acct_amount_filled has no missing vals
assert_all_are_not_na(accounts_clean$acct_amount_filled)
```

### `The End`
