---
title: "Record Linkage"
author: "Mohamad Osman"
date: '2022-06-27'
output: rmarkdown::github_document
---

# Section 04: Record Linkage

### **`01-Small distance, small difference`**

-   Calculate the Damerau-Levenshtein distance between `"las angelos"` and `"los angeles"`.

```{r}
library("stringdist")
```

```{r}
# Calculate Damerau-Levenshtein distance
stringdist("las angelos", "los angeles", method = "dl")
```

-   Calculate the Longest Common Substring (LCS) distance between `"las angelos"` and `"los angeles"`.

```{r}
# Calculate LCS distance
stringdist("las angelos", "los angeles", method = "lcs")

```

-   Calculate the Jaccard distance between `"las angelos"` and `"los angeles"`.

```{r}
# Calculate Jaccard distance
stringdist("las angelos", "los angeles", method = "jaccard")
```

Why is the LCS distance higher than the Damerau-Levenshtein distance between `"las angelos"` and `"los angeles"`?

\- LCS distance only uses insertion and deletion, so it takes more operations to change a string to another.

### **`02-Fixing typos with string distance`**

```{r}
# Load packages 
library(dplyr)
library(tidyverse)
library(fuzzyjoin)
```

```{r}

city_actual <- c("new york", "los angeles", "atlanta", 
                  "san francisco", "las vegas")
cities <- data.frame(city_actual)
cities
```

```{r}

file_path <- file.path("..", "00_Datasets", "zagat.rds")
zagat <- readRDS(file_path)
head(zagat, 3)

```

-   Count the number of each variation of `city` name in `zagat`.

```{r}
# Count the number of each city variation
zagat %>%
  count(city)
```

-   Left join `zagat` and `cities` based on string distance using the `city` and `city_actual` columns.

-   Select the `name`, `city`, and `city_actual` columns.

```{r}
# Count the number of each city variation
zagat %>%
  count(city)

# Join zagat and cities and look at results
zagat %>%
  # Left join based on stringdist using city and city_actual cols
  stringdist_left_join(cities, by = c("city" = "city_actual")) %>%
  # Select the name, city, and city_actual cols
  select(name, city, city_actual)
```

### **03-Pair blocking**

-   Load the `reclin` package.

-   Generate all possible pairs of records between the `zagat` and `fodors` datasets.

```{r}
fodors_file <- file.path("..", "00_datasets", "fodors.rds")
fodors <- readRDS(fodors_file)
head(fodors)
```

```{r}
# Load reclin
library(reclin)

# Generate all possible pairs
pair_blocking(zagat, fodors)
```

-   Use pair blocking to generate only pairs that have matching values in the `city` column.

```{r}
# Load reclin
library(reclin)

# Generate pairs with same city
pair_blocking(zagat, fodors, blocking_var = "city")
```

-   Compare pairs by `name` using `lcs()` distance

```{r}
# Generate pairs
pair_blocking(zagat, fodors, blocking_var = "city") %>%
  # Compare pairs by name using lcs()
  compare_pairs(by = "name",
      default_comparator = lcs())
```

-   Compare pairs by `name`, `phone`, and `addr` using `jaro_winkler()`.

```{r}
# Generate pairs
pair_blocking(zagat, fodors, blocking_var = "city") %>%
  # Compare pairs by name, phone, addr
  compare_pairs(c("name", "phone", "addr"),
     default_comparator =  jaro_winkler())
```

### **`04-Score then select or select then score?`**

#### The record linkage process into the correct order: 

-   `clean the datasets.`

-   `Generate pairs of records.`

-   `Compare separate columns of each pair.`

-   `Score pairs using summing or probability.`

-   `Select pairs that are matches based on their score.`

-   `Link the datasets together.`

### **05- Putting it together**

During this chapter, you've cleaned up the `city` column of `zagat` using string similarity, as well as generated and compared pairs of restaurants from `zagat` and `fodors`. The end is near - all that's left to do is score and select pairs and link the data together, and you'll be able to begin your analysis in no time!

`reclin` and `dplyr` are loaded and `zagat` and `fodors` are available.

-   Score the pairs of records probabilistically.

```{r}
# Create pairs
pair_blocking(zagat, fodors, blocking_var = "city") %>%
  # Compare pairs
  compare_pairs(by = c("name", "addr"), default_comparator = jaro_winkler())%>%
  score_simsum()%>%

  # Score pairs
  score_problink()
```

-   Select the pairs that are considered matches.

-   Link the two data frames together.

```{r}
# Create pairs
pair_blocking(zagat, fodors, blocking_var = "city") %>%
  # Compare pairs
  compare_pairs(by = c("name", "addr"), default_comparator = jaro_winkler()) %>%
  # Score pairs
  score_problink() %>%
  # Select pairs
  select_n_to_m() %>%
  link()
```

### `The End`
