---
title: "Record Linkage"
author: "Mohamad Osman"
date: '2022-06-27'
output: rmarkdown::github_document
---

# Section 04: Record Linkage

### **`01-Small distance, small difference`**

-   Calculate the Damerau-Levenshtein distance between `"las angelos"` and `"los angeles"`.

```{r}
library("stringdist")
```

```{r}
# Calculate Damerau-Levenshtein distance
stringdist("las angelos", "los angeles", method = "dl")
```

-   Calculate the Longest Common Substring (LCS) distance between `"las angelos"` and `"los angeles"`.

```{r}
# Calculate LCS distance
stringdist("las angelos", "los angeles", method = "lcs")

```

-   Calculate the Jaccard distance between `"las angelos"` and `"los angeles"`.

```{r}
# Calculate Jaccard distance
stringdist("las angelos", "los angeles", method = "jaccard")
```

Why is the LCS distance higher than the Damerau-Levenshtein distance between `"las angelos"` and `"los angeles"`?

- LCS distance only uses insertion and deletion, so it takes more operations to change a string to another.

\

\


