---
title: "Unsupervised learning in R"
author: "Mohamad Osman"
date: '2022-08-03'
output: rmarkdown::github_document
---

# Section 01: Unsupervised learning in R

### **`01-Identify clustering problems`**

Which of the following are clustering problems?

1.  Determining how many features it takes to describe most of the variability in data

2.  Determining the natural groupings of houses for sale based on size, number of bedrooms, etc.

3.  Visualizing 13 dimensional data (data with 13 features)

4.  Determining if there are common patterns in the demographics of people at a commerce site

5.  Predicting if someone will click on a web advertisement

Ans:

-   1, 3, and 5

-   2 and 3

-   1, 2, and 4

-   2 and 4 ✔️

-   All 5 are clustering problems

### 
**`02-k-means clustering`**

We have created some two-dimensional data and stored it in a variable called `x` in your workspace. The scatter plot on the right is a visual representation of the data.

In this exercise, your task is to create a k-means model of the `x` data using 3 clusters, then to look at the structure of the resulting model using the [**`summary()`**](https://www.rdocumentation.org/packages/base/topics/summary) function.

-   Fit a k-means model to `x` using 3 centers and run the k-means algorithm 20 times. Store the result in `km.out`.

-   Inspect the result with the `summary()` function.

```{r}
library(dplyr)
library(ggplot2)

x_file <- file.path("..", "00_Datasets", "x_k_means.txt")
x <- read.delim(x_file) %>%
     select(-id)

head(x)
```

```{r}
# Create the k-means model: km.out
km.out <- kmeans(x, centers = 3, nstart = 20)

# Inspect the result
summary(km.out)
```

### **`03-Results of kmeans()`**

The `kmeans()` function produces several outputs. In the video, we discussed one output of modeling, the cluster membership.

In this exercise, you will access the `cluster` component directly. This is useful anytime you need the cluster membership for each observation of the data used to build the clustering model. A future exercise will show an example of how this cluster membership might be used to help communicate the results of k-means modeling.

k-means models also have a print method to give a human friendly output of basic modeling results. This is available by using `print()` or simply typing the name of the model.

```{r}
# Print the cluster membership component of the model
km.out
print("-----------------------")
# Print the km.out object
print(km.out)
```

### **`04-Visualizing and interpreting results of kmeans()`**

One of the more intuitive ways to interpret the results of k-means models is by plotting the data as a scatter plot and using color to label the samples' cluster membership. In this exercise, you will use the standard `plot()` function to accomplish this.

To create a scatter plot, you can pass data with two features (i.e., columns) to `plot()` with an extra argument `col = km.out$cluster`, which sets the color of each point in the scatter plot according to its cluster membership.

`x` and `km.out` are available in your workspace. Using the `plot()` function to create a scatter plot of data `x`:

-   Color the dots on the scatterplot by setting the `col` argument to the `cluster` component in `km.out`.

-   Title the plot `"k-means with 3 clusters"` using the `main` argument to `plot()`.

-   Ensure there are no axis labels by specifying `""` for both the `xlab` and `ylab` arguments to `plot()`.

```{r}
# Scatter plot of x
plot(x, col = km.out$cluster,
     main = "k-means with 3 clusters",
     xlab = "",
     ylab = "")
```
