---
title: "Exploring Categorical Data"
author: "Mohamad Osman"
date: '2022-07-08'
output: rmarkdown::github_document
---

# Section 01: **Exploring Categorical Data**

### **`01-Contingency table review`**

In this chapter you'll continue working with the `comics` dataset introduced in the video. This is a collection of characteristics on all of the superheroes created by Marvel and DC comics in the last 80 years.

Let's start by creating a contingency table, which is a useful way to represent the total counts of observations that fall into each combination of the levels of categorical variables.

```{r}
library(readr)

url_csv <- "https://assets.datacamp.com/production/repositories/537/datasets/8860af2c0ef67fc77a8c704a73bbb93a395debcf/comics.csv"

comics <- read_csv(url_csv, col_types = "ffffffffiff")
```

```{r}
# Print the comics data
head(comics,3)

# Check levels of align
levels(comics$align)

# Check the levels of gender
levels(comics$gender)

# Create a 2-way contingency table
tab <- table(comics$align, comics$gender)
tab
```

### **`02-Dropping levels`**

The contingency table from the last exercise revealed that there are some levels that have very low counts. To simplify the analysis, it often helps to drop such levels.

In R, this requires two steps: first filtering out any rows with the levels that have very low counts, then removing these levels from the factor variable with `droplevels()`. This is because the `droplevels()` function would keep levels that have just 1 or 2 counts; it only drops levels that don't exist in a dataset.

The contingency table from the last exercise is available in your workspace as `tab`

-   Load the `dplyr` package.

-   Print `tab` to find out which level of `align` has the fewest total entries.

-   Use `filter()` to filter out all rows of `comics` with that level, then drop the unused level with `droplevels()`. Save the simplified dataset as `comics_filtered`

```{r}
# Load dplyr
library(dplyr)

# Print tab
print(tab)

# Remove align level
comics_filtered <- comics %>%
  filter(align != "Reformed Criminals") %>%
  droplevels()

# See the result
head(comics_filtered)

comics <- comics_filtered 
```

### **`03-Side-by-side bar charts`**

While a contingency table represents the counts numerically, it's often more useful to represent them graphically.

Here you'll construct two side-by-side bar charts of the `comics` data. This shows that there can often be two or more options for presenting the same data. Passing the argument `position = "dodge"` to `geom_bar()` says that you want a side-by-side (i.e. not stacked) bar chart.

-   Load the `ggplot2` package.

-   Create a side-by-side bar chart with `align` on the x-axis and `gender` as the `fill` aesthetic.

-   Create another side-by-side bar chart with `gender` on the x-axis and `align` as the `fill` aesthetic. Rotate the axis labels 90 degrees to help readability.

```{r}
# Load ggplot2
library(ggplot2)

# Create side-by-side bar chart of gender by alignment
ggplot(comics, aes(x = align, fill = gender)) + 
  geom_bar(position = "dodge")

# Create side-by-side bar chart of alignment by gender
ggplot(comics, aes(x = gender, fill = align)) + 
  geom_bar(position = "dodge") +
  theme(axis.text.x = element_text(angle = 90))
```

### **`04-Conditional proportions`**

The following code generates tables of joint and conditional proportions, respectively:

```{r}
tab <- table(comics$align, comics$gender)
options(scipen = 999, digits = 3) # Print fewer digits
prop.table(tab)     # Joint proportions
prop.table(tab, 1)  # Conditional on rows
prop.table(tab, 2)  # Conditional on columns
```

Go ahead and run it in the console. Approximately what proportion of all female characters are good?

Ans. 0.29 -\> 29%

### **`05-Counts vs. proportions (2)`**

Bar charts can tell dramatically different stories depending on whether they represent counts or proportions and, if proportions, what the proportions are conditioned on. To demonstrate this difference, you'll construct two bar charts in this exercise: one of counts and one of proportions.

-   Create a stacked bar chart of `gender` *counts* with `align` on the x-axis.

-   Create a stacked bar chart of `gender` *proportions* with `align` on the x-axis by setting the `position` argument to `geom_bar()` equal to `"fill"`.

```{r}
# Plot of gender by align
ggplot(comics, aes(x = align, fill = gender)) +
  geom_bar()
  
# Plot proportion of gender, conditional on align
ggplot(comics, aes(x = align, fill = gender)) + 
  geom_bar(position = "fill") +
  ylab("proportion")
```

### `06-Marginal bar chart`

If you are interested in the distribution of alignment of *all* superheroes, it makes sense to construct a bar chart for just that single variable.

You can improve the interpretability of the plot, though, by implementing some sensible ordering. Superheroes that are `"Neutral"` show an alignment between `"Good"` and `"Bad"`, so it makes sense to put that bar in the middle.

-   Reorder the levels of `align` using the [**`factor()`**](https://www.rdocumentation.org/packages/base/topics/factor) function so that printing them reads `"Bad"`, `"Neutral"`, then `"Good"`.

-   Create a bar chart of counts of the `align` variable.

```{r}
# Change the order of the levels in align
comics$align <- factor(comics$align, 
                       levels = c("Bad", "Neutral", "Good"))

# Create plot of align
ggplot(comics, aes(x = align)) + 
  geom_bar()
```

### **`07-Conditional bar chart`**

Now, if you want to break down the distribution of alignment based on gender, you're looking for conditional distributions.

You could make these by creating multiple filtered datasets (one for each gender) or by faceting the plot of alignment based on gender. As a point of comparison, we've provided your plot of the marginal distribution of alignment from the last exercise.

Create a bar chart of `align` faceted by `gender`.

```{r}
# Plot of alignment broken down by gender
ggplot(comics, aes(x = align)) + 
  geom_bar() +
  facet_wrap(~ gender)
```

### `08-Improve pie chart`

The pie chart is a very common way to represent the distribution of a single categorical variable, but they can be more difficult to interpret than bar charts.

This is a pie chart of a dataset called `pies` that contains the favorite pie flavors of 98 people. Improve the representation of these data by constructing a *bar chart* that is ordered in descending order of count.

```{r}
pies_path <- file.path("..", "00_Datasets", "flavor.txt")
pies <- read.delim(pies_path)
head(pies)
```

-   Use the code provided to reorder the levels of `flavor` so that they're in descending order by count.

-   Create a bar chart of `flavor` and orient the labels vertically so that they're easier to read. The default coloring may look drab by comparison, so change the `fill` of the bars to `"chartreuse"`.

```{r}
# Put levels of flavor in descending order
lev <- c("apple", "key lime", "boston creme", "blueberry", "cherry", "pumpkin", "strawberry")
pies$flavor <- factor(pies$flavor, levels = lev)

# Create bar chart of flavor
ggplot(pies, aes(x = flavor)) + 
  geom_bar(fill = "chartreuse") + 
  theme(axis.text.x = element_text(angle = 90))
```

### 
`The End`






