---
title: "Exploring Numerical Data"
author: "Mohamad Osman"
date: '2022-07-09'
output: rmarkdown::github_document
---

# **`Section 02: Exploring Numerical Data`**

### **`01-Faceted histogram`**

In this chapter, you'll be working with the `cars` dataset, which records characteristics on all of the new models of cars for sale in the US in a certain year. You will investigate the distribution of mileage across a categorical variable, but before you get there, you'll want to familiarize yourself with the dataset.

The `cars` dataset has been loaded in your workspace.

```{r}
library(readr)
library(dplyr)

url_csv <- "https://assets.datacamp.com/production/repositories/537/datasets/c0366d5da5ee8dce49919a5443685cf2e50c6a96/cars04.csv"

cars <- read_csv(url_csv)
head(cars,3)
```

-   Load the `ggplot2` package.

-   View the size of the data and the variable types using `str()`.

-   Plot a histogram of `city_mpg` faceted by `suv`, a logical variable indicating whether the car is an SUV or not.

```{r}
# Load package
library(ggplot2)

# Learn data structure
str(cars)

# Create faceted histogram
ggplot(cars, aes(x = city_mpg)) +
  geom_histogram() +
  facet_wrap(~ suv)
```

### **`02-Boxplots and density plots`**

The mileage of a car tends to be associated with the size of its engine (as measured by the number of cylinders). To explore the relationship between these two variables, you could stick to using histograms, but in this exercise you'll try your hand at two alternatives: the box plot and the density plot.

A quick look at `unique(cars$ncyl)` shows that there are more possible levels of `ncyl` than you might think. Here, restrict your attention to the most common levels.

-   Filter `cars` to include only cars with 4, 6, or 8 cylinders and save the result as `common_cyl`. The `%in%` operator may prove useful here.

-   Create side-by-side box plots of `city_mpg` separated out by `ncyl`.

-   Create overlaid density plots of `city_mpg` colored by `ncyl`.

```{r}
# Filter cars with 4, 6, 8 cylinders
common_cyl <- filter(cars, ncyl %in% c(4,6,8))

# Create box plots of city mpg by ncyl
ggplot(common_cyl, aes(x = as.factor(ncyl), y = city_mpg)) +
  geom_boxplot()

# Create overlaid density plots for same data
ggplot(common_cyl, aes(x = city_mpg, fill = as.factor(ncyl))) +
  geom_density(alpha = .3)
```

-   The highest mileage cars have 4 cylinders.

-   The typical 4 cylinder car gets better mileage than the typical 6 cylinder car, which gets better mileage than the typical 8 cylinder car.

-   Most of the 4 cylinder cars get better mileage than even the most efficient 8 cylinder cars.

### 
**`03-Marginal and conditional histograms`**

Now, turn your attention to a new variable: `horsepwr`. The goal is to get a sense of the marginal distribution of this variable and then compare it to the distribution of horsepower conditional on the price of the car being less than \$25,000.

You'll be making two plots using the "data pipeline" paradigm, where you start with the raw data and end with the plot.

-   Create a histogram of the distribution of `horsepwr` across all cars and add an appropriate title. Start by piping in the raw dataset.

-   Create a second histogram of the distribution of horsepower, but only for those cars that have an `msrp` less than \$25,000. Keep the limits of the x-axis so that they're similar to that of the first plot, and add a descriptive title.

```{r}
# Create hist of horsepwr
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram() +
  ggtitle("The Distribution of cars in terms of hores power")

# Create hist of horsepwr for affordable cars
cars %>% 
  filter(msrp < 25000) %>%
  ggplot(aes(horsepwr)) +
  geom_histogram() +
  xlim(c(90, 550)) +
  ggtitle("The Distribution of Cars horse power with prices < 25000$ ")
```

### **`04-Three binwidths`**

Before you take these plots for granted, it's a good idea to see how things change when you alter the binwidth. The binwidth determines how smooth your distribution will appear: the smaller the binwidth, the more jagged your distribution becomes. It's good practice to consider several binwidths in order to detect different types of structure in your data.

Create the following three plots, adding a title to each to indicate the binwidth used:

-   A histogram of horsepower (i.e. `horsepwr`) with a `binwidth` of 3.

-   A second histogram of horsepower with a `binwidth` of 30.

-   A third histogram of horsepower with a `binwidth` of 60.

```{r}
# Create hist of horsepwr with binwidth of 3
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram(binwidth = 3) +
  ggtitle("The distribution of cars in terms of horse power (3 bins)")

# Create hist of horsepwr with binwidth of 30
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram(binwidth = 30) +
  ggtitle("The distribution of cars in terms of horse power (30 bins)")

# Create hist of horsepwr with binwidth of 60
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram(binwidth = 60) +
  ggtitle("The distribution of cars in terms of horse power (60 bins)")

```

### **`05-Three binwidths interpretation`**

What feature is present in Plot A that's not found in B or C?

-   There is a tendency for cars to have horsepower right at 200 or 300 horsepower.

### **`06-Box plots for outliers`**

In addition to indicating the center and spread of a distribution, a box plot provides a graphical means to detect outliers. You can apply this method to the `msrp` column (manufacturer's suggested retail price) to detect if there are unusually expensive or cheap cars.

-   Construct a box plot of `msrp`.

-   Exclude the largest 3-5 outliers by filtering the rows to retain cars less than \$100,000. Save this reduced dataset as `cars_no_out`.

-   Construct a similar box plot of `msrp` using this reduced dataset. Compare the two plots.

```{r}
# Construct box plot of msrp
cars %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()

# Exclude outliers from data
cars_no_out <- cars %>%
  filter(msrp < 100000)

# Construct box plot of msrp using the reduced dataset
cars_no_out %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()
```

### **`07-Plot selection`**

Consider two other columns in the `cars` dataset: `city_mpg` and `width`. Which is the most appropriate plot for displaying the important features of their distributions? Remember, both density plots and box plots display the central tendency and spread of the data, but the box plot is more robust to outliers.

Use density plots or box plots to construct the following visualizations. For each variable, try both plots and submit the one that is better at capturing the important structure.

-   Display the distribution of `city_mpg`.

-   Display the distribution of `width`.

```{r}
# Create plot of city_mpg
cars %>%
  ggplot(aes(x = 1, y = city_mpg)) +
  geom_boxplot()

# Create plot of width
cars %>% 
  ggplot(aes(x = width)) +
  geom_density()
```

### **`08-3 variable plot`**

Faceting is a valuable technique for looking at several conditional distributions at the same time. If the faceted distributions are laid out in a grid, you can consider the association between a variable and two others, one on the rows of the grid and the other on the columns.

```{r}
# Facet hists using hwy mileage and ncyl
common_cyl %>%
  ggplot(aes(x = hwy_mpg)) +
  geom_histogram() +
  facet_grid(ncyl ~ suv) +
  ggtitle("Facet Histograms of hwy mile/gallon using ncyl and suv")
```

-   Across both SUVs and non-SUVs, mileage tends to decrease as the number of cylinders increases.

### **`The End`**
