---
title: "Case Study"
author: "Mohamad Osman"
date: '2022-07-10'
output: rmarkdown::github_document
---

# **`Section 04: Case Study`**

### **`01-Spam and num_char`**

Is there an association between spam and the length of an email? You could imagine a story either way:

-   *Spam is more likely to be a short message tempting me to click on a link*, or

-   *My normal email is likely shorter since I exchange brief emails with my friends all the time*.

Here, you'll use the `email` dataset to settle that question. Begin by bringing up the help file and learning about all the variables with `?email`.

As you explore the association between spam and the length of an email, use this opportunity to try out linking a `dplyr` chain with the layers in a `ggplot2` object

Using the `email` dataset

-   Load the packages `ggplot2`, `dplyr`, and `openintro`.

-   Compute appropriate measures of the center and spread of `num_char` for both spam and not-spam using `group_by()` and `summarize()`. No need to name the new columns created by `summarize()`.

-   Construct side-by-side box plots to visualize the association between the same two variables. It will be useful to `mutate()` a new column containing a log-transformed version of `num_char`.

```{r}
# Load packages
library(dplyr)
library(ggplot2)
library(openintro)
```

```{r}
# Compute summary statistics
email %>%
  group_by(spam) %>%
  summarize(median(num_char),
            IQR(num_char))

# Create plot
email %>%
  mutate(log_num_char = log(num_char)) %>%
  ggplot(aes(x = spam, y = log_num_char)) +
  geom_boxplot()
```

### **`02-Spam and !!!`**

Let's look at a more obvious indicator of spam: exclamation marks. `exclaim_mess` contains the number of exclamation marks in each message. Using summary statistics and visualization, see if there is a relationship between this variable and whether or not a message is spam.

Experiment with different types of plots until you find one that is the most informative. Recall that you've seen:

-   Side-by-side box plots

-   Faceted histograms

-   Overlaid density plots

The `email` dataset is still available in your workspace.

-   Calculate appropriate measures of the center and spread of `exclaim_mess` for both spam and not-spam using `group_by()` and `summarize()`.

-   Construct an appropriate plot to visualize the association between the same two variables, adding in a log-transformation step if necessary.

-   If you decide to use a log transformation, remember that `log(0)` is `-Inf` in R, which isn't a very useful value! You can get around this by adding a small number (like `0.01`) to the quantity inside the `log()` function. This way, your value is never zero. This small shift to the right won't affect your results.

```{r}
# Compute center and spread for exclaim_mess by spam
email %>%
    group_by(spam) %>%
    summarise(IQR(exclaim_mess),
              median(exclaim_mess))


# Create plot for spam and exclaim_mess
email %>% 
    group_by(spam )%>%
    ggplot(aes(x = log(exclaim_mess, 0.01))) +
    geom_histogram() + 
    facet_wrap(~ spam)

```

### **`03-Collapsing levels`**

If it was difficult to work with the heavy skew of `exclaim_mess`, the number of images attached to each email (`image`) poses even more of a challenge. Run the following code at the console to get a sense of its distribution:

    table(email$image)

Recall that this tabulates the number of cases in each category (so there were 3811 emails with 0 images, for example). Given the very low counts at the higher number of images, let's collapse `image` into a categorical variable that indicates whether or not the email had at least one image. In this exercise, you'll create this new variable and explore its association with spam.

Starting with `email`, form a continuous chain that links together the following tasks:

-   Create a new variable called `has_image` that is `TRUE` where the number of images is greater than zero and `FALSE` otherwise.

-   Create an appropriate plot with `email` to visualize the relationship between `has_image` and `spam`.

```{r}
# Create plot of proportion of spam by image
email %>%
  mutate(has_image = image > 0) %>%
  ggplot(aes(x = has_image, fill = spam)) +
  geom_bar(position = "fill")
```

### **`04-Data Integrity`**

In the process of exploring a dataset, you'll sometimes come across something that will lead you to question how the data were compiled. For example, the variable `num_char` contains the number of characters in the email, in thousands, so it could take decimal values, but it certainly shouldn't take negative values.

You can formulate a test to ensure this variable is behaving as we expect:

    email$num_char < 0

If you run this code at the console, you'll get a long vector of logical values indicating for each case in the dataset whether that condition is `TRUE`. Here, the first 1000 values all appear to be `FALSE`. To verify that *all* of the cases indeed have non-negative values for `num_char`, we can take the *sum* of this vector:

    sum(email$num_char < 0)

This is a handy shortcut. When you do arithmetic on logical values, R treats `TRUE` as `1` and `FALSE` as `0`. Since the sum over the whole vector is zero, you learn that every case in the dataset took a value of `FALSE` in the test. That is, the `num_char` column is behaving as we expect and taking only non-negative values.

Consider the variables `image` and `attach`. You can read about them with `?email`, but the help file is ambiguous: do attached images count as attached files in this dataset?

Design a simple test to determine if images count as attached files. This involves creating a logical condition to compare the values of the two variables, then using `sum()` to assess every case in the dataset. Recall that the logical operators are `<` for *less than*, `<=` for *less than or equal to*, `>` for *greater than*, `>=` for *greater than or equal to*, and `==` for *equal to*.

```{r}
# Test if images count as attachments
sum(email$image > email$attach)


```

### **`05-Answering questions with chains`**

When you have a specific question about a dataset, you can find your way to an answer by carefully constructing the appropriate chain of R code. For example, consider the following question: "Within non-spam emails, is the typical length of emails shorter for those that were sent to multiple people?"

This can be answered with the following chain:

    email %>%
       filter(spam == "not-spam") %>%
       group_by(to_multiple) %>%
       summarize(median(num_char))

The code makes it clear that you are using `num_char` to measure the length of an email and `median()` as the measure of what is typical. If you run this code, you'll learn that the answer to the question is "yes": the typical length of non-spam sent to multiple people is a bit lower than those sent to only one person.

This chain concluded with summary statistics, but others might end in a plot; it all depends on the question that you're trying to answer.

Build a chain to answer each of the following questions, both about the variable `dollar`.

-   For emails containing the word "dollar", does the typical spam email contain a greater number of occurrences of the word than the typical non-spam email? Create a summary statistic that answers this question.

-   If you encounter an email with greater than 10 occurrences of the word `dollar`, is it more likely to be spam or not-spam? Create a bar chart that answers this question.

```{r}
# Question 1
email %>%
  filter(dollar > 0) %>%
  group_by(spam) %>%
  summarize(median(dollar))

# Question 2
email %>%
  filter(dollar > 10) %>%
  ggplot(aes(x = spam)) +
  geom_bar()
```

### `06-What's in a number?`

Turn your attention to the variable called `number`. Read more about it by pulling up the help file with `?email`.

To explore the association between this variable and `spam`, select and construct an informative plot. For illustrating relationships between categorical variables, you've seen

-   Faceted bar charts

-   Side-by-side bar charts

-   Stacked and normalized bar charts.

Let's practice constructing a faceted bar chart.

-   Reorder the levels of `number` so that they preserve the natural ordering of `"none"`, then `"small"`, then `"big"`, saving to a `number_reordered` column.

-   Construct a faceted bar chart of the association between `number` and `spam`

```{r}
# Reorder levels
email$number_reordered <- factor(email$number, 
                                levels = c("none", "small", "big"))

# Construct plot of number_reordered
ggplot(email, aes(x = number_reordered)) +
  geom_bar() +
  facet_wrap(~ spam)
```

### `The End`
