---
title: "Importing data from databases"
author: "Mohamad Osman"
date: '2022-06-22'
output: rmarkdown::github_document
---

# Section 01: Importing data from databases

### **`01-Establish a connection`**

The first step to import data from a SQL database is creating a connection to it. As Filip explained, you need different packages depending on the database you want to connect to. All of these packages do this in a uniform way, as specified in the `DBI` package.

[**`dbConnect()`**](https://www.rdocumentation.org/packages/DBI/functions/dbConnect) creates a connection between your R session and a SQL database. The first argument has to be a `DBIdriver` object, that specifies how connections are made and how data is mapped between R and the database. Specifically for MySQL databases, you can build such a driver with [**`RMySQL::MySQL()`**](https://www.rdocumentation.org/packages/RMySQL/functions/MySQLDriver-class).

If the MySQL database is a remote database hosted on a server, you'll also have to specify the following arguments in [**`dbConnect()`**](https://www.rdocumentation.org/packages/DBI/functions/dbConnect): `dbname`, `host`, `port`, `user` and `password`. Most of these details have already been provided.

-   Load the `DBI` library, which is already installed on DataCamp's servers.

-   Edit the [**`dbConnect()`**](https://www.rdocumentation.org/packages/DBI/functions/dbConnect) call to connect to the MySQL database. Change the `port` argument (`3306`) and `user` argument (`"student"`).

```{r}
# install.packages("RMySQL")
```

```{r}
# Load the DBI package
library(DBI)

# Edit dbConnect() call
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "tweater", 
                 host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                 port = 3306,
                 user = "student",
                 password = "datacamp")

con

# disconnect to the database
dbDisconnect(con)
```

### **`02-List the database tables`**

-   Add code to create a vector `tables`, that contains the tables in the tweater database. You can connect to this database through the `con` object.

-   Display the structure of `tables`; what's the class of this vector?

```{r}
# Load the DBI package
library(DBI)

# Connect to the MySQL database: con
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "tweater", 
                 host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                 port = 3306,
                 user = "student",
                 password = "datacamp")

# Build a vector of table names: tables
tables <- dbListTables(con)

# Display structure of tables
str(tables)

# disconnect to the database
dbDisconnect(con)
```

### **`03-Import users`**

There are three tables: **`users`**, **`tweats`**, and **`comments`** that have relations among them. Which ones, you ask? You'll discover in a moment!

Let's start by importing the data on the users into your R session. You do this with the [**`dbReadTable()`**](https://www.rdocumentation.org/packages/DBI/functions/dbReadTable) function. Simply pass it the connection object (`con`), followed by the name of the table you want to import. The resulting object is a standard R data frame.

-   Add code that imports the `"users"` table from the tweater database and store the resulting data frame as `users`.

-   Print the `users` data frame.

```{r}
# Load the DBI package
library(DBI)

# Connect to the MySQL database: con
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "tweater", 
                 host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                 port = 3306,
                 user = "student",
                 password = "datacamp")

# Import the users table from tweater: users
users <- dbReadTable(con, "users")

# Print users
users

# disconnect to the database
dbDisconnect(con)
```

### **`04-Import all tables`**

Next to the `users`, we're also interested in the `tweats` and `comments` tables. However, separate `dbReadTable()` calls for each and every one of the tables in your database would mean a lot of code duplication. Remember about the `lapply()` function? You can use it again here! A connection is already coded for you, as well as a vector `table_names`, containing the names of all the tables in the database.

-   Finish the `lapply()` function to import the `users`, `tweats` and `comments` tables in a single call. The result, a list of data frames, will be stored in the variable `tables`.

-   Print `tables` to check if you got it right.

```{r}
# Load the DBI package
library(DBI)

# Connect to the MySQL database: con
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "tweater", 
                 host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                 port = 3306,
                 user = "student",
                 password = "datacamp")

# Get table names
table_names <- dbListTables(con)

# Import all tables
tables <- lapply(table_names, dbReadTable, conn = con)

# Print out tables
tables

# disconnect to the database
dbDisconnect(con)
```

### **`The End`**

\
