---
title: "Importing data from the web (Part 1)"
author: "Mohamad Osman"
date: '2022-06-22'
output: rmarkdown::github_document
---

# Section 03: **Importing data from the web (Part 1)**

### **`01-Import flat files from the web`**

-   Load the `readr` package. It's already installed on DataCamp's servers.

-   Use `url_csv` to read in the `.csv` file it is pointing to. Use the [**`read_csv()`**](https://cran.r-project.org/web/packages/readr/readr.pdf) function. The `.csv` contains column names in the first row. Save the resulting data frame as `pools`.

-   Similarly, use `url_delim` to read in the online `.txt` file. Use the [**`read_tsv()`**](https://cran.r-project.org/web/packages/readr/readr.pdf) function and store the result as `potatoes`.

-   Print `pools` and `potatoes`. Looks correct?

```{r}
# Load the readr package
library(readr)

# Import the csv file: pools
url_csv <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/swimming_pools.csv"
pools <- read_csv(url_csv) 

# Import the txt file: potatoes
url_delim <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/potatoes.txt"
potatoes <- read_tsv(url_delim)

# Print pools and potatoes
pools
potatoes
```

### **`02-Secure importing`**

In the previous exercises, you have been working with URLs that all start with `http://`. There is, however, a safer alternative to HTTP, namely HTTPS, which stands for HyperText Transfer Protocol Secure. Just remember this: HTTPS is relatively safe, HTTP is not.

Luckily for us, you can use the standard importing functions with `https://` connections since R version 3.2.2.

-   Take a look at the URL in `url_csv`. It uses a secure connection, `https://`.

-   Use [**`read.csv()`**](https://www.rdocumentation.org/packages/utils/functions/read.table) to import the file at `url_csv`. The `.csv` file it is referring to contains column names in the first row. Call it `pools1`.

-   Load the `readr` package. It's already installed on DataCamp's servers.

-   Use [**`read_csv()`**](https://cran.r-project.org/web/packages/readr/readr.pdf) to read in the same `.csv` file in `url_csv`. Call it `pools2`.

-   Print out the structure of `pools1` and `pools2`. Looks like the importing went equally well as with a normal `http` connection!

```{r}
# https URL to the swimming_pools csv file.
url_csv <- "https://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/swimming_pools.csv"

# Import the file using read.csv(): pools1
pools1 <- read.csv(url_csv)

# Load the readr package
library(readr)

# Import the file using read_csv(): pools2
pools2 <- read_csv(url_csv)

# Print the structure of pools1 and pools2
str(pools1)
str(pools2)
```

### **`03-Import Excel files from the web`**

-   Load the `readxl` and `gdata` packages. They are already installed on DataCamp's servers.

-   Import the `.xls` file located at the URL `url_xls` using [**`read.xls()`**](https://www.rdocumentation.org/packages/gdata/functions/read.xls) from `gdata`. Store the resulting data frame as `excel_gdata`.

-   You can not use [**`read_excel()`**](https://cran.r-project.org/web/packages/readxl/readxl.pdf) directly with a URL. Complete the following instructions to work around this problem:

-   Use [**`download.file()`**](https://www.rdocumentation.org/packages/utils/functions/download.file) to download the `.xls` file behind the URL and store it locally as `"local_latitude.xls"`.

-   Call [**`read_excel()`**](https://cran.r-project.org/web/packages/readxl/readxl.pdf) to import the local file, `"local_latitude.xls"`. Name the resulting data frame `excel_readxl`.

```{r}
# Load the readxl and gdata package
library(readxl)
library(gdata)

# Specification of url: url_xls
url_xls <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/latitude.xls"

# Import the .xls file with gdata: excel_gdata
excel_gdata <- gdata::read.xls(url_xls)

# Download file behind URL, name it local_latitude.xls
dest_path <- file.path("..", "00_Datasets", "local_latitude.xls")
download.file(url_xls, dest_path, mode="wb")


# Import the local .xls file with readxl: excel_readxl
excel_readxl <- readxl::read_excel(dest_path)
excel_readxl
```

### **`04-Downloading any file, secure or not`**

In the previous exercise you've seen how you can read excel files on the web using the `read_excel` package by first downloading the file with the [**`download.file()`**](https://www.rdocumentation.org/packages/utils/functions/download.file) function.

There's more: with [**`download.file()`**](https://www.rdocumentation.org/packages/utils/functions/download.file) you can download any kind of file from the web, using HTTP and HTTPS: images, executable files, but also `.RData` files. An `RData` file is very efficient format to store R data.

You can load data from an `RData` file using the [**`load()`**](https://www.rdocumentation.org/packages/base/functions/load) function, but this function does not accept a URL string as an argument. In this exercise, you'll first download the `RData` file securely, and then import the local data file.

-   Take a look at the URL in `url_rdata`. It uses a secure connection, `https://`. This URL points to an `RData` file containing a data frame with some metrics on different kinds of wine.

-   Download the file at `url_rdata` using [**`download.file()`**](https://www.rdocumentation.org/packages/utils/functions/download.file). Call the file `"wine_local.RData"` in your working directory.

-   Load the file you created, `wine_local.RData`, using the [**`load()`**](https://www.rdocumentation.org/packages/base/functions/load) function. It takes one argument, the path to the file, which is just the filename in our case. After running this command, the variable `wine` will automatically be available in your workspace.

Print out the [**`summary()`**](https://www.rdocumentation.org/packages/base/functions/summary) of the `wine` dataset.

```{r}
# https URL to the wine RData file.
url_rdata <- "https://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/wine.RData"

# Download the wine file to your working directory
dest_path <- file.path("..", "00_Datasets", "wine_local.RData")

download.file(url_rdata, dest_path)

# Load the wine data into your workspace using load()
load(dest_path)

# Print out the summary of the wine data
summary(wine)
```

### **`05-HTTP? httr! (1)`**

`httr` provides a convenient function, [**`GET()`**](https://www.rdocumentation.org/packages/httr/functions/GET) to execute this GET request. The result is a `response` object, that provides easy access to the status code, content-type and, of course, the actual content.

-   Load the `httr` package. It's already installed on DataCamp's servers.

-   Use [**`GET()`**](https://www.rdocumentation.org/packages/httr/functions/GET) to get the URL stored in `url`. Store the result of this `GET()` call as `resp`.

-   Print the `resp` object. What information does it contain?

-   Get the content of `resp` using [**`content()`**](https://www.rdocumentation.org/packages/httr/functions/content) and set the `as` argument to `"raw"`. Assign the resulting vector to `raw_content`.

-   Print the first values in `raw_content` with [**`head()`**](https://www.rdocumentation.org/packages/utils/functions/head).

```{r}
# Load the httr package
library(httr)

# Get the url, save response to resp
url <- "http://www.example.com/"
resp <- GET(url)

# Print resp
print(resp)

# Get the raw content of resp: raw_content
raw_content <- content(resp, as = "raw")

# Print the head of raw_content
print(raw_content)
```

### **`06-HTTP? httr! (2)`**

-   Use [**`GET()`**](https://www.rdocumentation.org/packages/httr/functions/GET) to get the `url` that has already been specified in the sample code. Store the response as `resp`.

-   Print `resp`. What is the content-type?

-   Use [**`content()`**](https://www.rdocumentation.org/packages/httr/functions/content) to get the content of `resp`. Set the `as` argument to `"text"`. Simply print out the result. What do you see?

-   Use [**`content()`**](https://www.rdocumentation.org/packages/httr/functions/content) to get the content of `resp`, but this time do not specify a second argument. R figures out automatically that you're dealing with a JSON, and converts the JSON to a named R list.

```{r}
# httr is already loaded

# Get the url
url <- "http://www.omdbapi.com/?apikey=72bc447a&t=Annie+Hall&y=&plot=short&r=json"
resp <- GET(url)

# Print resp
print(resp)

# Print content of resp as text
print(content(resp, as = "text"))

# Print content of resp
print(content(resp))

```

### `The End`
