---
title: "Importing data from the web (Part 2)"
author: "Mohamad Osman"
date: '2022-06-22'
output: rmarkdown::github_document
---

# **Section 04: Importing data from the web (Part 2)**

### **`01-From JSON to R`** 

In the simplest setting, [**`fromJSON()`**](https://www.rdocumentation.org/packages/jsonlite/functions/fromJSON) can convert character strings that represent JSON data into a nicely structured R list. Give it a try!

-   Load the `jsonlite` package. It's already installed on DataCamp's servers.

-   `wine_json` represents a JSON. Use `fromJSON()` to convert it to a list, named `wine`.

-   Display the structure of `wine`

```{r}
# Load the jsonlite package
library(jsonlite)

# wine_json is a JSON
wine_json <- '{"name":"Chateau Migraine", "year":1997, "alcohol_pct":12.4, "color":"red", "awarded":false}'

# Convert wine_json into a list: wine
wine <- fromJSON(wine_json)

# Print structure of wine
str(wine)
```

### **`02-Quandl API`**

-   `quandl_url` represents a URL. Use `fromJSON()` directly on this URL and store the result in `quandl_data`.

-   Display the structure of `quandl_data`.

```{r}
# jsonlite is preloaded

# Definition of quandl_url
quandl_url <- "https://www.quandl.com/api/v3/datasets/WIKI/FB/data.json?auth_token=i83asDsiWUUyfoypkgMz"

# Import Quandl data: quandl_data
quandl_data <- fromJSON(quandl_url)

# Print structure of quandl_data
str(quandl_data)
```

### **`03-OMDb API`**

-   Two URLs are included in the sample code, as well as a `fromJSON()` call to build `sw4`. Add a similar call to build `sw3`.

-   Print out the element named `Title` of both `sw4` and `sw3`. You can use the `$` operator. What movies are we dealing with here?

-   Write an expression that evaluates to `TRUE` if `sw4` was released later than `sw3`. This information is stored in the `Year` element of the named lists.

```{r}
# The package jsonlite is already loaded

# Definition of the URLs
url_sw4 <- "http://www.omdbapi.com/?apikey=72bc447a&i=tt0076759&r=json"
url_sw3 <- "http://www.omdbapi.com/?apikey=72bc447a&i=tt0121766&r=json"

# Import two URLs with fromJSON(): sw4 and sw3
sw4 <- fromJSON(url_sw4)
sw3 <- fromJSON(url_sw3)

# Print out the Title element of both lists
print(sw4$Title)
print(sw3$Title)


# Is the release year of sw4 later than sw3?
sw4$Year > sw3$Year
```

### **`04-JSON practice (1)`**

-   Change the assignment of `json1` such that the R vector after conversion contains the numbers 1 up to 6, in ascending order. Next, call [**`fromJSON()`**](https://www.rdocumentation.org/packages/jsonlite/functions/fromJSON) on `json1`.

-   Adapt the code for `json2` such that it's converted to a named list with two elements: `a`, containing the numbers 1, 2 and 3 and `b`, containing the numbers 4, 5 and 6. Next, call [**`fromJSON()`**](https://www.rdocumentation.org/packages/jsonlite/functions/fromJSON) on `json2`.

```{r}
# jsonlite is already loaded

# Challenge 1
json1 <- '[1, 2, 3, 4, 5, 6]'
fromJSON(json1)

# Challenge 2
json2 <- '{"a": [1, 2, 3], "b": [4, 5, 6]}'
fromJSON(json2)
```

### **`05-JSON practice (2)`**

-   Remove characters from `json1` to build a 2 by 2 matrix containing only 1, 2, 3 and 4. Call [**`fromJSON()`**](https://www.rdocumentation.org/packages/jsonlite/functions/fromJSON) on `json1`.

-   Add characters to `json2` such that the data frame in which the json is converted contains an additional observation in the last row. For this observations, `a` equals 5 and `b` equals 6. Call [**`fromJSON()`**](https://www.rdocumentation.org/packages/jsonlite/functions/fromJSON) one last time, on `json2`.

```{r}
# jsonlite is already loaded

# Challenge 1
json1 <- '[[1, 2], [3, 4]]'
fromJSON(json1)

# Challenge 2
json2 <- '[{"a": 1, "b": 2}, {"a": 3, "b": 4}, {"a": 5, "b": 6}]'
fromJSON(json2)
```

### **`06-toJSON()`**

-   Use a function of the `utils` package to import the `.csv` file directly from the URL specified in `url_csv`. Save the resulting data frame as `water`. Make sure that strings are *not* imported as factors.

-   Convert the data frame `water` to a JSON. Call the resulting object `water_json`.

-   Print out `water_json`.

```{r}
# jsonlite is already loaded

# URL pointing to the .csv file
url_csv <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/water.csv"

# Import the .csv file located at url_csv
water <- read.csv(url_csv)
water
# Convert the data file according to the requirements
water_json <- toJSON(water)

# Print out water_json
water_json
```

### **`07-Minify and prettify`**

JSONs can come in different formats. Take these two JSONs, that are in fact exactly the same: the first one is in a minified format, the second one is in a pretty format with indentation, whitespace and new lines:

    # Mini
    {"a":1,"b":2,"c":{"x":5,"y":6}}

    # Pretty
    {
      "a": 1,
      "b": 2,
      "c": {
        "x": 5,
        "y": 6
      }
    }

Unless you're a computer, you surely prefer the second version. However, the standard form that [**`toJSON()`**](https://www.rdocumentation.org/packages/jsonlite/functions/fromJSON) returns, is the minified version, as it is more concise. You can adapt this behavior by setting the `pretty` argument inside [**`toJSON()`**](https://www.rdocumentation.org/packages/jsonlite/functions/fromJSON) to `TRUE`. If you already have a JSON string, you can use [**`prettify()`**](https://www.rdocumentation.org/packages/jsonlite/functions/prettify) or [**`minify()`**](https://www.rdocumentation.org/packages/jsonlite/functions/prettify) to make the JSON pretty or as concise as possible.

-   Convert the `mtcars` dataset, which is available in R by default, to a *pretty* `JSON`. Call the resulting JSON `pretty_json`.

-   Print out `pretty_json`. Can you understand the output easily?

-   Convert `pretty_json` to a minimal version using [**`minify()`**](https://www.rdocumentation.org/packages/jsonlite/functions/prettify). Store this version under a new variable, `mini_json`.

-   Print out `mini_json`. Which version do you prefer, the pretty one or the minified one?

```{r}
# jsonlite is already loaded

# Convert mtcars to a pretty JSON: pretty_json
mtcars_sliced <- mtcars[20:30,]

pretty_json <- toJSON(mtcars_sliced, pretty = TRUE)

# Print pretty_json
pretty_json

# Minify pretty_json: mini_json
mini_json <- minify(pretty_json)

# Print mini_json
mini_json
```

### **`The End`**
