---
title: "Importing data from statistical software packages"
author: "Mohamad Osman"
date: '2022-06-23'
output: rmarkdown::github_document
---

# **Section 05: Importing data from statistical software packages**

### **`01-Import SAS data with haven`**

-   Load the `haven` package; it's already installed on DataCamp's servers.

-   Import the data file `"sales.sas7bdat"`. Call the imported data frame `sales`.

-   Display the structure of `sales` with [**`str()`**](https://www.rdocumentation.org/packages/utils/functions/str). Some columns represent categorical variables, so they should be factors.

```{r}
# Load the haven package
library(haven)

# Import sales.sas7bdat: sales
dest_path <- file.path("..", "00_Datasets", "sales.sas7bdat")


sales <- read_sas(dest_path)

# Display the structure of sales
str(sales)
```

### **`02-Import STATA data with haven`**

-   Import the data file directly from the URL using [**`read_dta()`**](http://cran.r-project.org/web/packages/haven/haven.pdf), and store it as `sugar`.

-   Print out the structure of `sugar`. The `Date` column has class `labelled`.

-   Convert the values in the `Date` column of `sugar` to dates, using `as.Date(as_factor(___))`.

-   Print out the structure of `sugar` once more. Looks better now?

```{r}
# haven is already loaded

# Import the data from the URL: sugar
sugar <- read_dta("http://assets.datacamp.com/production/course_1478/datasets/trade.dta")

# Structure of sugar
str(sugar)

# Convert values in Date column to dates
sugar$Date <- as.Date(as_factor(sugar$Date))

# Structure of sugar again
str(sugar)
```

### **`03-Import SPSS data with haven`**

-   Use [**`read_sav()`**](http://cran.r-project.org/web/packages/haven/haven.pdf) to import the SPSS data in `"person.sav"`. Name the imported data frame `traits`.

-   `traits` contains several missing values, or `NA`s. Run [**`summary()`**](https://www.rdocumentation.org/packages/base/functions/summary) on it to find out how many `NA`s are contained in each variable.

-   Print out a subset of those individuals that scored high on Extroversion *and* on Agreeableness, i.e. scoring higher than 40 on each of these two categories. You can use `subset()` for this.

```{r}
# haven is already loaded

# Import person.sav: traits
dest_path <- file.path("..", "00_Datasets", "person.sav")

traits <- read_sav(dest_path)

# Summarize traits
summary(traits)

# Print out a subset
subset(traits, Extroversion > 40 & Agreeableness > 40)
```

### **`04-Factorize, round two`**

-   Import the SPSS data straight from the URL and store the resulting data frame as `work`.

-   Display the summary of the `GENDER` column of `work`. This information doesn't give you a lot of useful information, right?

-   Convert the `GENDER` column in `work` to a factor, the class to denote categorical variables in R. Use [**`as_factor()`**](http://cran.r-project.org/web/packages/haven/haven.pdf).

-   Once again display the summary of the `GENDER` column. This time, the printout makes much more sense.

```{r}
# haven is already loaded

# Import SPSS data from the URL: work
work <- read_sav("http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/employee.sav")

# Display summary of work$GENDER
summary(work$GENDER)

# Convert work$GENDER to a factor
work$GENDER <- as_factor(work$GENDER)

# Display summary of work$GENDER again
summary(work$GENDER)
```

### **`05-Import STATA data with foreign (1)`**

The `foreign` package offers a simple function to import and read *STATA* data: [**`read.dta()`**](https://www.rdocumentation.org/packages/foreign/functions/read.dta).

In this exercise, you will import data on the US presidential elections in the year 2000. The data in `florida.dta` contains the total numbers of votes for each of the four candidates as well as the total number of votes per election area in the state of Florida (Source: [**Florida Department of State**](https://results.elections.myflorida.com/)). The file is available in your working directory, you can download it [**here**](http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/florida.dta) if you want to experiment some more.

-   Load the `foreign` package; it's already installed on DataCamp's servers.

-   Import the data on the elections in Florida, `"florida.dta"`, and name the resulting data frame `florida`. Use [**`read.dta()`**](https://www.rdocumentation.org/packages/foreign/functions/read.dta) without specifying extra arguments.

-   Check out the last 6 observations of `florida` with `tail()`

```{r}
# Load the foreign package
library(foreign)

# File path 
dest_path <- file.path("..", "00_Datasets", "florida.dta")


# Import florida.dta and name the resulting data frame florida
florida <- read.dta(dest_path)

# Check tail() of florida
tail(florida)
```

### **`06-Import STATA data with foreign (2)`**

The arguments you will use most often are `convert.dates`, `convert.factors`, `missing.type` and `convert.underscore`. Their meaning is pretty straightforward, as Filip explained in the video. It's all about correctly converting STATA data to standard R data structures. Type `?read.dta` to find out about about the default values.

-   Specify the path to the file using [**`file.path()`**](https://www.rdocumentation.org/packages/base/functions/file.path). Call it `path`. Remember the `"edequality.dta"` file is located in the `"worldbank"` folder.

-   Use the `path` variable to import the data file in three different ways; each time show its structure with [**`str()`**](https://www.rdocumentation.org/packages/utils/functions/str):

-   `edu_equal_1`: By passing only the file `path` to [**`read.dta()`**](https://www.rdocumentation.org/packages/foreign/functions/read.dta).

-   `edu_equal_2`: By passing the file `path`, and setting `convert.factors` to `FALSE`.

-   `edu_equal_3`: By passing the file `path`, and setting `convert.underscore` to `TRUE`.

```{r}
# foreign is already loaded

# Specify the file path using file.path(): path
path <- file.path("..", "00_Datasets", "edequality.dta")
str(path)

# Create and print structure of edu_equal_1
edu_equal_1 <- read.dta(path)
str(edu_equal_1)

# Create and print structure of edu_equal_2
edu_equal_2 <- read.dta(path, convert.factors = FALSE)
str(edu_equal_2)

# Create and print structure of edu_equal_3
edu_equal_3 <- read.dta(path, convert.underscore = TRUE)
str(edu_equal_3)
```

### **`07-Import SPSS data with foreign (1)`**

All great things come in pairs. Where `foreign` provided [**`read.dta()`**](https://www.rdocumentation.org/packages/foreign/functions/read.dta) to read Stata data, there's also [**`read.spss()`**](https://www.rdocumentation.org/packages/foreign/functions/read.spss) to read SPSS data files. To get a data frame, make sure to set `to.data.frame = TRUE` inside [**`read.spss()`**](https://www.rdocumentation.org/packages/foreign/functions/read.spss).

In this exercise, you'll be working with socio-economic variables from different countries (Source: [**Quantative Data Analysis in Education**](http://cw.routledge.com/textbooks/9780415372985/resources/datasets.asp)). The SPSS data is in a file called `international.sav`, which is in your working directory. You can also download it [**here**](http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/international.sav) if you want to play around with it some more.

-   Import the data file `"international.sav"` and have R convert it to a data frame. Store this data frame as `demo`.

-   Create a boxplot of the `gdp` variable of `demo`.

```{r}
# foreign is already loaded

# Import international.sav as a data frame: demo
path <- file.path("..", "00_Datasets", "international.sav")

demo <- read.spss(path, to.data.frame = TRUE)

# Create boxplot of gdp variable of demo
boxplot(demo$gdp)
```

### 
**`08-Import SPSS data with foreign (2)`**

-   Import the data file `"international.sav"` as a data frame, `demo_1`.

-   Print the first few rows of `demo_1` using the [**`head()`**](https://www.rdocumentation.org/packages/utils/functions/head) function.

-   Import the data file `"international.sav"` as a data frame, `demo_2`, but this time in a way such that variables with value labels are *not* converted to R factors.

-   Again, print the first few rows of `demo_2`. Can you tell the difference between the two data frames?

```{r}
# foreign is already loaded

# Import international.sav as demo_1
demo_1 <- read.spss(path, to.data.frame = TRUE)

# Print out the head of demo_1
head(demo_1)

# Import international.sav as demo_2
demo_2 <- read.spss(path, to.data.frame = TRUE, use.value.labels = FALSE)

# Print out the head of demo_2
head(demo_2)
```

### `The End`
  
