---
title: "Simple Linear Regression"
author: "Mohamad Osman"
date: '2022-07-19'
output: rmarkdown::github_document
---

# Section 01: Simple Linear Regression

### `01-Visualizing two variables`

```{r}
library(dplyr)
library(ggplot2)
library(fst)

file_path <- file.path("..", "00_Datasets", "taiwan_real_estate.fst")
taiwan_real_estate <- read_fst(file_path)
taiwan_real_estate$house_age_years <- as.factor(taiwan_real_estate$house_age_years)
head(taiwan_real_estate, 3)
```

Before you can run any statistical models, it's usually a good idea to visualize your dataset. Here, we'll look at the relationship between house price per area and the number of nearby convenience stores, using the Taiwan real estate dataset.

One challenge in this dataset is that the number of convenience stores contains integer data, causing points to overlap. To solve this, you will make the points transparent.

`taiwan_real_estate` is available, `ggplot2` is loaded, and its black and white theme has been set.

-   Using `taiwan_real_estate`, draw a scatter plot of `price_twd_msq` (y-axis) versus `n_convenience` (x-axis).

```{r}
# Draw a scatter plot of n_convenience vs. price_twd_msq
ggplot(taiwan_real_estate, aes(n_convenience, price_twd_msq)) + 

geom_point()
```

-   Update the plot to make the points 50% transparent by setting `alpha` to `0.5`.

-   Update the plot by adding a trend line, calculated using a linear regression. You can omit the confidence ribbon.

```{r}
# Add a linear trend line without a confidence ribbon
ggplot(taiwan_real_estate, aes(n_convenience, price_twd_msq)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", 
              se = FALSE)
```

### `02-Linear regression with lm()`

While ggplot can display a linear regression trend line using `geom_smooth()`, it doesn't give you access to the intercept and slope as variables, or allow you to work with the model results as variables. That means that sometimes you'll need to run a linear regression yourself.

Time to run your first model!

`taiwan_real_estate` is available. *TWD* is an abbreviation for Taiwan dollars.

-   Run a linear regression with `price_twd_msq` as the response variable, `n_convenience` as the explanatory variable, and `taiwan_real_estate` as the dataset.

```{r}
# Run a linear regression of price_twd_msq vs. n_convenience
lm(price_twd_msq ~ n_convenience, data = taiwan_real_estate)
```

-   The model had an `(Intercept)` coefficient of `8.2242`. What does this mean?

    `Answer:` On average, a house with zero convenience stores nearby had a price of 8.2242 TWD per square meter.

-   The model had an `n_convenience` coefficient of `0.7981`. What does this mean?
    `Answer:` If you increase the number of nearby convenience stores by one, then the expected increase in house price is `0.7981` TWD per square meter.

### **`03-Visualizing numeric vs. categorical`**

If the explanatory variable is categorical, the scatter plot that you used before to visualize the data doesn't make sense. Instead, a good option is to draw a histogram for each category.

The Taiwan real estate dataset has a categorical variable in the form of the age of each house. The ages have been split into 3 groups: 0 to 15 years, 15 to 30 years, and 30 to 45 years.

`taiwan_real_estate` is available and `ggplot2` is loaded.

-   Using `taiwan_real_estate`, plot a histogram of `price_twd_msq` with `10` bins.

-   Facet the plot by `house_age_years` to give 3 panels.

```{r}
# Using taiwan_real_estate, plot price_twd_msq
ggplot(taiwan_real_estate, aes(price_twd_msq)) +
  # Make it a histogram with 10 bins
  geom_histogram(bins = 10) +
  # Facet the plot so each house age group gets its own panel
  facet_wrap(~ house_age_years)
```

### `04-Calculating means by category`

A good way to explore categorical variables is to calculate summary statistics such as the mean for each category. Here, you'll look at grouped means for the house prices in the Taiwan real estate dataset.

`taiwan_real_estate` is available and `dplyr` is loaded.

-   Group `taiwan_real_estate` by `house_age_years`.

-   Summarize to calculate the mean `price_twd_msq` for each group, naming the column `mean_by_group`.

-   Assign the result to `summary_stats` and *look at the numbers*.

```{r}
summary_stats <- taiwan_real_estate %>% 
  # Group by house age
  group_by(house_age_years) %>% 
  # Summarize to calculate the mean house price/area
  summarise(mean_by_group = mean(price_twd_msq))

# See the result
summary_stats
```

### **`05-lm() with a categorical explanatory variable`**

Linear regressions also work with categorical explanatory variables. In this case, the code to run the model is the same, but the coefficients returned by the model are different. Here you'll run a linear regression on the Taiwan real estate dataset.

`taiwan_real_estate` is available.

-   Run a linear regression with `price_twd_msq` as the response variable, `house_age_years` as the explanatory variable, and `taiwan_real_estate` as the dataset. Assign to `mdl_price_vs_age`.

```{r}
# Update the model formula to remove the intercept
mdl_price_vs_age_no_intercept <- lm(
  price_twd_msq ~ house_age_years, 
  data = taiwan_real_estate
)

# See the result
mdl_price_vs_age_no_intercept
```

-   Update the model formula so that no intercept is included in the model. Assign to `mdl_price_vs_age_no_intercept`.

```{r}
# Update the model formula to remove the intercept
mdl_price_vs_age_no_intercept <- lm(
  price_twd_msq ~ house_age_years + 0, 
  data = taiwan_real_estate
)

# See the result
mdl_price_vs_age_no_intercept
```

### `The End` 







\

\

\

\



\

\


