---
title: "Parallel Slopes"
author: "Mohamad Osman"
date: '2022-07-21'
output: rmarkdown::github_document
---

# Section 01: **Parallel Slopes**

### `01-Fitting a parallel slopes linear regression`

In [**Introduction to Regression in R**](https://learn.datacamp.com/courses/introduction-to-regression-in-r), you learned to fit linear regression models with a single explanatory variable. In many cases, using only one explanatory variable limits the accuracy of predictions. That means that to truly master linear regression, you need to be able to include multiple explanatory variables.

The case when there is one numeric explanatory variable and one categorical explanatory variable is sometimes called a "parallel slopes" linear regression due to the shape of the predictions---more on that in the next exercise.

Here, you'll revisit the Taiwan real estate dataset. Recall the meaning of each variable.

```{r}
library(ggplot2)
library(fst)
library(dplyr)
library(broom)

file_path <- file.path("..", "00_Datasets", "taiwan_real_estate.fst")
taiwan_real_estate <- read_fst(file_path)
taiwan_real_estate$house_age_years <- as.factor(taiwan_real_estate$house_age_years)
head(taiwan_real_estate, 3)
```

| Variable                | Meaning                                                             |
|:-------------------|:---------------------------------------------------|
| `dist_to_mrt_station_m` | Distance to nearest MRT metro station, in meters.                   |
| `n_convenience`         | No. of convenience stores in walking distance.                      |
| `house_age_years`       | The age of the house, in years, in 3 groups.                        |
| `price_twd_msq`         | House price per unit area, in New Taiwan dollars per meter squared. |

`taiwan_real_estate` is available.

-   Using the `taiwan_real_estate` dataset, model the house price (in TWD per square meter) versus the number of nearby convenience stores.

```{r}
# Fit a linear regr'n of price_twd_msq vs. n_convenience
mdl_price_vs_conv <- lm(price_twd_msq ~ n_convenience, data = taiwan_real_estate)

# See the result
mdl_price_vs_conv
```

-   Model the house price (in TWD per square meter) versus the house age (in years). Don't include an intercept term.

```{r}
# Fit a linear regr'n of price_twd_msq vs. house_age_years, no intercept
mdl_price_vs_age <- lm(price_twd_msq ~ house_age_years + 0, data = taiwan_real_estate)

# See the result
mdl_price_vs_age
```

-   Model the house price (in TWD per square meter) versus the number of nearby convenience stores plus the house age (in years). Don't include an intercept term.

```{r}
# Fit a linear regr'n of price_twd_msq vs. n_convenience plus house_age_years, no intercept
mdl_price_vs_both <- lm(price_twd_msq ~ n_convenience + house_age_years + 0, data = taiwan_real_estate)

# See the result
mdl_price_vs_both
```

### **`02-Interpreting parallel slopes coefficients`**

For linear regression with a single numeric explanatory variable, there is an intercept coefficient and a slope coefficient. For linear regression with a single categorical explanatory variable, there is an intercept coefficient for each category.

In the "parallel slopes" case, where you have a numeric and a categorical explanatory variable, what do the coefficients mean?

`taiwan_real_estate` and `mdl_price_vs_both` are available.

Look at the coefficients of `mdl_price_vs_both`. What is the meaning of the `n_convenience` coefficient?

```{r}
coefficients(mdl_price_vs_both)
```

For each additional nearby convenience store, the expected house price, in TWD per square meter, increases by `0.79`.

-   What is the meaning of the `"0 to 15 years"` coefficient?

What is the meaning of the `"0 to 15 years"` coefficient?

-   For a house aged 0 to 15 years with zero nearby convenience stores, the expected house price is `9.41` TWD per square meter.

### **`03-Visualizing each explanatory variable`**

Being able to see the predictions made by a model makes it easier to understand. In the case where there is only one explanatory variable, ggplot lets you do this without any manual calculation or messing about.

To visualize the relationship between a numeric explanatory variable and the numeric response, you can draw a scatter plot with a linear trend line.

To visualize the relationship between a categorical explanatory variable and the numeric response, you can draw a box plot.

`taiwan_real_estate` is available and `ggplot2` is loaded.

-   Using the `taiwan_real_estate` dataset, plot the house price versus the number of nearby convenience stores.

-   Make it a scatter plot.

-   Add a smooth linear regression trend line without a standard error ribbon.

```{r}
# Using taiwan_real_estate, plot price_twd_msq vs. n_convenience
ggplot(taiwan_real_estate, aes(n_convenience ,price_twd_msq)) +
  # Add a point layer
  geom_point() +
  # Add a smooth trend line using linear regr'n, no ribbon
  geom_smooth(method = "lm", se = FALSE)
```

-   Using the `taiwan_real_estate` dataset, plot the house price versus the house age.

-   Make it a box plot.

```{r}
# Using taiwan_real_estate, plot price_twd_msq vs. house_age_years
ggplot(taiwan_real_estate, aes(house_age_years, price_twd_msq)) +
  # Add a box plot layer
  geom_boxplot()
```

### **`04-Visualizing parallel slopes`**

The two plots in the previous exercise gave very different predictions: one gave a predicted response that increased linearly with a numeric variable; the other gave a fixed response for each category. The only sensible way to reconcile these two conflicting predictions is to incorporate both explanatory variables in the model at once.

When it comes to a linear regression model with a numeric and a categorical explanatory variable, `ggplot2` doesn't have an easy, "out of the box" way to show the predictions. Fortunately, the `moderndive` package includes an extra geom, `geom_parallel_slopes()` to make it simple.

`taiwan_real_estate` is available; `ggplot2` and `moderndive` are loaded.

-   Using the `taiwan_real_estate` dataset, plot house prices versus the number of nearby convenience stores, colored by house age.

-   Make it a scatter plot.

-   Add parallel slopes, without a standard error ribbon.

```{r}
library(moderndive)

# Using taiwan_real_estate, plot price_twd_msq vs. n_convenience colored by house_age_years
ggplot(taiwan_real_estate, aes(n_convenience , price_twd_msq,
color = house_age_years)) +
  # Add a point layer
  geom_point() +
  # Add parallel slopes, no ribbon
  geom_parallel_slopes(se = FALSE)
```

Parallel slope paradise! The "parallel slope" model name comes from the fact that the prediction for each category is a slope, and all those slopes are parallel.

### **`05-Predicting with a parallel slopes model`**

While ggplot can automatically show you model predictions, in order to get those values to program with, you'll need to do the calculations yourself.

Just as with the case of a single explanatory variable, the workflow has two steps: create a data frame of explanatory variables, then add a column of predictions. To make sure you've got the right answer, you can add your predictions to the ggplot with the `geom_parallel_slopes()` lines.

`taiwan_real_estate` and `mdl_price_vs_both` are available; `dplyr`, `tidyr`, and `ggplot2` are loaded.

note: `expand_grid` is deprecated and it is replaced by `expand.grid`.

```{r}
# Make a grid of explanatory data
# note expand_grid is replaced by expand.replace
explanatory_data <- expand.grid(
  # Set n_convenience to zero to ten
  n_convenience = seq(0,10,1),
  # Set house_age_years to the unique values of that variable
  house_age_years = unique(taiwan_real_estate$house_age_years)
)

# See the result
explanatory_data
```

-   Add a column to the `explanatory_data` named for the response variable, assigning to `prediction_data`.

-   The response column contain predictions made using `mdl_price_vs_both` and `explanatory_data`.

```{r}
# From previous step
explanatory_data <- expand.grid(
  n_convenience = 0:10,
  house_age_years = unique(taiwan_real_estate$house_age_years)
)

# Add predictions to the data frame
prediction_data <- explanatory_data %>% 
  mutate(price_twd_msq = predict(mdl_price_vs_both, explanatory_data))

# See the result
prediction_data
```

-   Update the plot to add a point layer of predictions. Use the `prediction_data`, set the point size to 5, and the point shape to 15.

```{r}
# From previous steps
explanatory_data <- expand.grid(
  n_convenience = 0:10,
  house_age_years = unique(taiwan_real_estate$house_age_years)
)
prediction_data <- explanatory_data %>% 
  mutate(
    price_twd_msq = predict(mdl_price_vs_both, explanatory_data)
  )

taiwan_real_estate %>% 
  ggplot(aes(n_convenience, price_twd_msq, color = house_age_years)) +
  geom_point() +
  geom_parallel_slopes(se = FALSE) +
  # Add points using prediction_data, with size 5 and shape 15
  geom_point(data = prediction_data, size = 5, shape = 15)
```

### **`06-Manually calculating predictions`**

As with simple linear regression, you can manually calculate the predictions from the model coefficients. The only change for the parallel slopes case is that the intercept is different for each category of the categorical explanatory variable. That means you need to consider the case when each each category occurs separately.

`taiwan_real_estate`, `mdl_price_vs_both`, and `explanatory_data` are available; `dplyr` is loaded.

-   Get the coefficients from `mdl_price_vs_both`, assigning to `coeffs`.

-   Assign each of the elements of `coeffs` to the appropriate variable.

```{r}
# Get the coefficients from mdl_price_vs_both
coeffs <- coefficients(mdl_price_vs_both)

# Extract the slope coefficient
slope <- coeffs[1]

# Extract the intercept coefficient for 0 to 15
intercept_0_15 <- coeffs[2]

# Extract the intercept coefficient for 15 to 30
intercept_15_30 <- coeffs[3]

# Extract the intercept coefficient for 30 to 45
intercept_30_45 <- coeffs[4]
```

Add columns to `explanatory_data`.

-   To choose the `intercept`, in the case when `house_age_years` is `"0 to 15"`, choose `intercept_0_15`. In the case when `house_age_years` is `"15 to 30"`, choose `intercept_15_30`. Do likewise for `"30 to 45"`

-   Manually calculate the predictions as the intercept plus the slope times `n_convenience`.

```{r}
# From previous step
coeffs <- coefficients(mdl_price_vs_both)
slope <- coeffs[1]
intercept_0_15 <- coeffs[2]
intercept_15_30 <- coeffs[3]
intercept_30_45 <- coeffs[4]

prediction_data <- explanatory_data %>% 
  mutate(
    # Consider the 3 cases to choose the intercept
    intercept = case_when(
        house_age_years == "0 to 15" ~ intercept_0_15,
        house_age_years == "15 to 30" ~ intercept_15_30, 
        house_age_years =="30 to 45" ~ intercept_30_45
    ),
    
    # Manually calculate the predictions
    price_twd_msq = intercept + slope * n_convenience
  )

# See the results
prediction_data
```

### **`07-Comparing coefficients of determination`**

Recall that the coefficient of determination is a measure of how well the linear regression line fits the observed values. An important motivation for including several explanatory variables in a linear regression is that you can improve the fit compared to considering only a single explanatory variable.

Here you'll compare the coefficient of determination for the three Taiwan house price models, to see which gives the best result.

`mdl_price_vs_conv`, `mdl_price_vs_age`, and `mdl_price_vs_both` are available; `dplyr` and `broom` are loaded.

-   Get the unadjusted and adjusted coefficients of determination for `mdl_price_vs_conv` by glancing at the model, then selecting the `r.squared` and `adj.r.squared` values.

-   Do the same for `mdl_price_vs_age` and `mdl_price_vs_both`.

```{r}
mdl_price_vs_conv %>% 
  # Get the model-level coefficients
  glance() %>% 
  # Select the coeffs of determination
  select(r.squared, adj.r.squared)
  
# Get the coeffs of determination for mdl_price_vs_age
mdl_price_vs_age %>% 
  glance() %>% 
  select(r.squared, adj.r.squared)

# Get the coeffs of determination for mdl_price_vs_both
mdl_price_vs_both %>% 
  glance() %>% 
  select(r.squared, adj.r.squared)
```

Which model does the adjusted coefficient of determination suggest gives a better fit?

-   `mdl_price_vs_both`.

### **`08-Comparing residual standard error`**

The other common metric for assessing model fit is the residual standard error (RSE), which measures the typical size of the residuals.

In the last exercise you saw how including both explanatory variables into the model increased the coefficient of determination. How do you think using both explanatory variables will change the RSE?

`mdl_price_vs_conv`, `mdl_price_vs_age`, and `mdl_price_vs_both` are available; `dplyr` and `broom` are loaded.

-   Get the residual standard error for `mdl_price_vs_conv` by glancing at the model, then pulling the `sigma` value.

-   Do the same for `mdl_price_vs_age`.

-   Do the same for `mdl_price_vs_both`.

```{r}
mdl_price_vs_conv %>% 
  # Get the model-level coefficients
  glance %>% 
  # Pull out the RSE
  pull(sigma)

# Get the RSE for mdl_price_vs_age
mdl_price_vs_age %>% 
  glance() %>%
  pull(sigma)


# Get the RSE for mdl_price_vs_both
mdl_price_vs_both %>% 
  glance() %>%
  pull(sigma)
```

-   Which model does the RSE suggest gives more accurate predictions?

    -   `mdl_price_vs_both`. with residual 3.21

### `The End`

\

\

\

\

\

\
