---
title: "01_Getting_Started_with_R_Markdown"
output:
  html_document: default
  pdf_document: default
date: '2022-06-19'
---

# Section 01: **Getting Started with R Markdown**

### `01-Creating your first R Markdown file`

-   Add the investment_annual_summary dataset to the code chunk in your report.

```{r data, include = FALSE}
library(readr)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
```

```{r}
investment_annual_summary
```

### `02-Adding code chunks to your file`

-   Add a new code chunk to the Markdown file, starting at line `14`.

-   Within the code chunk, type the names of the `investment_annual_summary` and `investment_services_projects` datasets to print them in the report.

```{r}
library(readr)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/78b002735b6f620df7f2767e63b76aaca317bf8d/investment_services_projects.csv")
```

```{r}
investment_annual_summary
investment_services_projects
```

### **`03-Adding sections to your report`**

-   Add a header called Datasets to line `14` of the report, using two hashes.

-   Add headers to lines `16` and `23` of the report to specify each of the dataset names, Investment Annual Summary and Investment Services Projects, using three hashes.

-   Add the dataset names to the sentences in lines `18` and `25` of the report and format them as inline code.

```{r data, include = FALSE}
library(readr)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/78b002735b6f620df7f2767e63b76aaca317bf8d/investment_services_projects.csv")
```

## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018.

```{r}
investment_annual_summary
```

### Investment Services Projects

The `investment_services_projects` dataset provides information about each investment project from the 2012 to 2018 fiscal years. Information listed includes the project name, company name, sector, project status, and investment amounts.

```{r}
investment_services_projects 
```

### **`04-Editing the YAML header`**

-   Add the `author` field to the YAML header beneath the title and add your name.

-   Add the `date` field below the author information using the `Sys.Date()` function.

---
title: "Investment Report"
author: "Mohamad Osman"
date: "`r Sys.Date()`"
output: html_document
---

### **`05-Formatting the date`**

-   Modify the date in the YAML header using the `Sys.time()` function so that the date displays as the decimal day, the full name of the month, and the four digit year, for example: 08 April 2020.

---
title: "Investment Report"
author: "Add your name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

### `The End`
