---
title: "Adding Analyses and Visualizations"
author: "Mohamad Osman"
date: '2022-06-19'
output: html_document
---

# **Section 02: Adding Analyses and Visualizations**

### **`01-Filtering for a specific country`**

-   Load the `dplyr` package in the first code chunk on line `9` right after the `readr` package on line `8`.

-   In the third code chunk, filter the `investment_services_projects` data to determine which projects occurred in Brazil.

-   Save the result as `brazil_investment_projects`.

-   Label the code chunk `brazil-investment-projects`.

---
title: "Investment Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r data, include = FALSE}
library(readr)
library(dplyr)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/bcb2e39ecbe521f4b414a21e35f7b8b5c50aec64/investment_services_projects.csv")
```

## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018.

```{r investment-annual-summary}
investment_annual_summary
```

### Investment Projects in Brazil

The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts.

```{r brazil-investment-projects}
brazil_investment_projects <- investment_services_projects %>%
  filter(country == "Brazil")

brazil_investment_projects
```

### **`02-Filtering for a specific year`**

-   In the fourth code chunk, create `brazil_investment_projects_2018` by filtering the `investment_services_projects` data for projects in Brazil with a `date_disclosed` in the 2018 fiscal year, which starts on July 1, 2017 and ends on June 30, 2018.

-   Label the code chunk `brazil-investment-projects-2018`.

-   Add a header to line `33` using three hashes to label the section `Investment Projects in Brazil in 2018`.

### Investment Projects in Brazil in 2018

```{r brazil-investment-projects-2018}
brazil_investment_projects_2018 <- investment_services_projects %>%
  filter(country == "Brazil",
         date_disclosed >= "2017-07-01",
         date_disclosed <= "2018-06-30") 

brazil_investment_projects_2018
```

### **`03-Referencing code results in the report`**

-   In the `brazil-investment-projects-2018` code chunk, create `brazil_investment_projects_2018_total` by summarizing the data that was filtered for projects in Brazil in the 2018 fiscal year to create a new column, `sum_total_investment`, that contains the sum of the `total_investment` amounts for each of the projects.

-   Add `Brazil` and `2018` to the sentence on line `47` to describe the information that was calculated.

-   Add a reference to `brazil_investment_projects_2018_total` in the sentence on line `47`, so that the total investment amount for all projects in Brazil in the 2018 fiscal year is included in the knit report.

```{r brazil-investment-projects-2018}
brazil_investment_projects_2018 <- investment_services_projects %>%
  filter(country == "Brazil",
         date_disclosed >= "2017-07-01",
         date_disclosed <= "2018-06-30") 

brazil_investment_projects_2018

brazil_investment_projects_2018_total <- brazil_investment_projects_2018 %>%
  summarize(sum_total_investment = sum(total_investment, na.rm = TRUE)) 
```

The total investment amount for all projects in Brazil in the 2018 fiscal year was `r brazil_investment_projects_2018_total` million dollars.

### **`04-Visualizing the Investment Annual Summary data`**

-   Load the `ggplot2` package in the first code chunk, after the `dplyr` package on line `10`.

-   In the `investment-annual-summary` code chunk, create a line plot using the `investment_annual_summary` data.

-   Plot the fiscal year (`fiscal_year`) on the x-axis, and investment amounts (`dollars_in_millions`) on the y-axis.

-   Color the plot by `region`, so that the plot in the final report displays each region as a different color in the line plot.

```{r data, include = FALSE}
library(readr)
library(dplyr)
library(ggplot2)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/bcb2e39ecbe521f4b414a21e35f7b8b5c50aec64/investment_services_projects.csv")
```

## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018.

```{r investment-annual-summary}
ggplot(investment_annual_summary, aes(x = fiscal_year, y = dollars_in_millions, color = region)) +
  geom_line() +
  labs(
    title = "Investment Annual Summary",
    x = "Fiscal Year",
    y = "Dollars in Millions"
  )
```

### Investment Projects in Brazil

The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts.

```{r brazil-investment-projects}
brazil_investment_projects <- investment_services_projects %>%
  filter(country == "Brazil") 
```

### Investment Projects in Brazil in 2018

```{r brazil-investment-projects-2018}
brazil_investment_projects_2018 <- investment_services_projects %>%
  filter(country == "Brazil",
         date_disclosed >= "2017-07-01",
         date_disclosed <= "2018-06-30")
```

### **`05-Visualizing all projects for one country`**

-   In the `brazil-investment-projects` code chunk, create a scatterplot of the `brazil_investment_projects` data.

-   Plot the date disclosed (`date_disclosed`) on the x-axis, and total IFC investment (`total_investment`) on the y-axis.

-   Color the plot by `status`, so that the plot includes information about the status of each project.

---
title: "Investment Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r data, include = FALSE}
library(readr)
library(dplyr)
library(ggplot2)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/bcb2e39ecbe521f4b414a21e35f7b8b5c50aec64/investment_services_projects.csv")
```

## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018.

```{r investment-annual-summary}
ggplot(investment_annual_summary, aes(x = fiscal_year, y = dollars_in_millions, color = region)) +
  geom_line() +
  labs(
    title = "Investment Annual Summary",
    x = "Fiscal Year",
    y = "Dollars in Millions"
  )
```

### Investment Projects in Brazil

The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts.

```{r brazil-investment-projects}
brazil_investment_projects <- investment_services_projects %>%
  filter(country == "Brazil") 

ggplot(brazil_investment_projects, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  )
```

### Investment Projects in Brazil in 2018

```{r brazil-investment-projects-2018}
brazil_investment_projects_2018 <- investment_services_projects %>%
  filter(country == "Brazil",
         date_disclosed >= "2017-07-01",
         date_disclosed <= "2018-06-30") 
```

### **06-Visualizing all projects for one country and year**

-   In the `brazil-investment-projects-2018` code chunk, create a scatterplot of the `brazil_investment_projects_2018` data.

-   Add the title "Investment Services Projects in Brazil in 2018" to the plot.

-   Label the x-axis "Date Disclosed" and the y-axis "Total IFC Investment in Dollars in Millions".

---
title: "Investment Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r data, include = FALSE}
library(readr)
library(dplyr)
library(ggplot2)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/bcb2e39ecbe521f4b414a21e35f7b8b5c50aec64/investment_services_projects.csv")
```

## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018.

```{r investment-annual-summary}
ggplot(investment_annual_summary, aes(x = fiscal_year, y = dollars_in_millions, color = region)) +
  geom_line() +
  labs(
    title = "Investment Annual Summary",
    x = "Fiscal Year",
    y = "Dollars in Millions"
  )
```

### Investment Projects in Brazil

The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts.

```{r brazil-investment-projects}
brazil_investment_projects <- investment_services_projects %>%
  filter(country == "Brazil") 

ggplot(brazil_investment_projects, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects in Brazil",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  )
```

### Investment Projects in Brazil in 2018

```{r brazil-investment-projects-2018}
brazil_investment_projects_2018 <- investment_services_projects %>%
  filter(country == "Brazil",
         date_disclosed >= "2017-07-01",
         date_disclosed <= "2018-06-30") 

ggplot(brazil_investment_projects_2018, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects in Brazil in 2018",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  ) 
```

### **`07-chunk options globally`**

-   Add to the `setup` chunk at the top of the report to align all figures in the center of the report.

-   Add another option to the `setup` chunk so that the output width of all figures are 80%

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = 'center', out.width = '80%', echo = TRUE)
```

### **`08-Setting chunk options locally`**

-   Add to the `investment-annual-summary` chunk so that the output dimensions of the resulting figure is 85%.

-   Add to the `brazil-investment-projects` and `brazil-investment-projects-2018` chunks so that the output dimensions of the figures are 95%.

---
title: "Investment Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = 'center', echo = TRUE)
```

```{r data, include = FALSE}
library(readr)
library(dplyr)
library(ggplot2)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/bcb2e39ecbe521f4b414a21e35f7b8b5c50aec64/investment_services_projects.csv")
```

## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018.

```{r investment-annual-summary, out.width = "85%"}
ggplot(investment_annual_summary, aes(x = fiscal_year, y = dollars_in_millions, color = region)) +
  geom_line() +
  labs(
    title = "Investment Annual Summary",
    x = "Fiscal Year",
    y = "Dollars in Millions"
  )
```

### Investment Projects in Brazil

The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts.

```{r brazil-investment-projects, out.width = "95%"}
brazil_investment_projects <- investment_services_projects %>%
  filter(country == "Brazil") 

ggplot(brazil_investment_projects, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects in Brazil",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  )
```

### Investment Projects in Brazil in 2018

```{r brazil-investment-projects-2018, out.width = "95%"}
brazil_investment_projects_2018 <- investment_services_projects %>%
  filter(country == "Brazil",
         date_disclosed >= "2017-07-01",
         date_disclosed <= "2018-06-30") 

ggplot(brazil_investment_projects_2018, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects in Brazil in 2018",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  ) 
```

### 

**`09-Adding figure captions`**

---
title: "Investment Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.align = 'center', echo = TRUE)
```

```{r data, include = FALSE}
library(readr)
library(dplyr)
library(ggplot2)

investment_annual_summary <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/d0251f26117bbcf0ea96ac276555b9003f4f7372/investment_annual_summary.csv")
investment_services_projects <- read_csv("https://assets.datacamp.com/production/repositories/5756/datasets/bcb2e39ecbe521f4b414a21e35f7b8b5c50aec64/investment_services_projects.csv")
```

## Datasets

### Investment Annual Summary

The `investment_annual_summary` dataset provides a summary of the dollars in millions provided to each region for each fiscal year, from 2012 to 2018.

```{r investment-annual-summary, out.width = '85%', fig.cap = 'Figure 1.1 The Investment Annual Summary for each region for 2012 to 2018'}
ggplot(investment_annual_summary, aes(x = fiscal_year, y = dollars_in_millions, color = region)) +
  geom_line() +
  labs(
    title = "Investment Annual Summary",
    x = "Fiscal Year",
    y = "Dollars in Millions"
  )
```

### Investment Projects in Brazil

The `investment_services_projects` dataset provides information about each investment project from 2012 to 2018. Information listed includes the project name, company name, sector, project status, and investment amounts.

```{r brazil-investment-projects, out.width = '95%', fig.cap = 'Figure 1.2 The Investment Services Projects in Brazil from 2012 to 2018'}
brazil_investment_projects <- investment_services_projects %>%
  filter(country == "Brazil") 

ggplot(brazil_investment_projects, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects in Brazil",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  )
```

### Investment Projects in Brazil in 2018

```{r brazil-investment-projects-2018, out.width = '95%', fig.cap = 'Figure 1.3 The Investment Services Projects in Brazil in 2018'}
brazil_investment_projects_2018 <- investment_services_projects %>%
  filter(country == "Brazil",
         date_disclosed >= "2017-07-01",
         date_disclosed <= "2018-06-30") 

ggplot(brazil_investment_projects_2018, aes(x = date_disclosed, y = total_investment, color = status)) +
  geom_point() +
  labs(
    title = "Investment Services Projects in Brazil in 2018",
    x = "Date Disclosed",
    y = "Total IFC Investment in Dollars in Millions"
  ) 
```
