---
title: "Dates and Times in R"
author: "Mohamad Osman"
date: '2022-06-28'
output: rmarkdown::github_document
---

# Section 01: Dates and Times in R

### **`01-Recognizing ISO 8601 dates`**

-   Which of the following is the correct way to specify the *4th of October 2004* according to `ISO 8601`?

    `Ans: 2004-10-04`

### **`02-Specifying dates`**

```{r}
x = "2013-04-03"
```

We've stored the string `"2013-04-03"` in a variable called `x`.

-   Use `str()` to look at the structure of `x` and confirm it's just a character string.

-   Convert `x` to a date using `as.Date()`.

-   Use `str()` to look at the structure of `x_date` and confirm it's a `Date`.

-   Now use `as.Date()` to store the date *April 10, 2014*.

```{r}
# The date R 3.0.0 was released
x <- "2013-04-03"

# Examine structure of x
str(x)

# Use as.Date() to interpret x as a date
x_date <- as.Date(x)

# Examine structure of x_date
str(x_date)

# Store April 10 2014 as a Date
april_10_2014 <- as.Date("2014-04-10")
april_10_2014
```

### **`03-Automatic import`**

-   Use `read_csv()` to read in the CSV file `rversions.csv` as `releases`.

-   Use `str()` to examine the structure of the `date` column. *Notice it's already a `Date` object*.

-   We've loaded `anytime` and created an object called `sep_10_2009`. Use the `anytime()` function to parse `sep_10_2009`.

```{r}
# Load the readr package
library(readr)

# Use read_csv() to import rversions.csv
rversions_path <- file.path("..", "00_Datasets", "rversions.csv")
releases <- read.csv(rversions_path)

# Examine the structure of the date column
str(releases$date)

# Load the anytime package
library(anytime)

# Various ways of writing Sep 10 2009
sep_10_2009 <- c("September 10 2009", "2009-09-10", "10 Sep 2009", "09-10-2009")

# Use anytime() to parse sep_10_2009
anytime(sep_10_2009)
```

### **`04-Plotting`**

-   Make a plot of releases over time by setting the `x` argument of the `aes()` function to the `date` column.

-   Zoom in to the period from 2010 to 2014 by specifying limits from `"2010-01-01"` to `"2014-01-01"`. *Notice these strings need to be wrapped in `as.Date()` to be interpreted as `Date` objects.*

-   Adjust the axis labeling by specifying `date_breaks` of `"10 years"` and `date_labels` of `"%Y"`.

```{r}
releases$date <- as.Date(releases$date)
str(releases)
```

```{r}
library(ggplot2)

# Set the x axis to the date column
ggplot(releases, aes(x = date, y = type)) +
  geom_line(aes(group = 1, color = factor(major)))

# Limit the axis to between 2010-01-01 and 2014-01-01
ggplot(releases, aes(x = date, y = type)) +
  geom_line(aes(group = 1, color = factor(major))) +
  xlim(as.Date("2010-01-01"), as.Date("2014-01-01"))

# Specify breaks every ten years and labels with "%Y"
ggplot(releases, aes(x = date, y = type)) +
  geom_line(aes(group = 1, color = factor(major))) +
  scale_x_date(date_breaks = "10 years", date_labels = "%Y")
```

### 05-**Arithmetic and logical operators**

-   Find the date of the most recent release by calling `max()` on the date column in `releases`.

-   Find the rows in `releases` that have the most recent date, by specifying the comparison `date == last_release_date` in [**`filter()`**](https://www.rdocumentation.org/packages/dplyr/versions/0.7.8/topics/filter)

-   Print `last_release` to see which release this was.

-   Calculate how long it has been since the most recent release by subtracting `last_release_date` from `Sys.Date()`

```{r}
library(dplyr)
```

```{r}
# Find the largest date
last_release_date <- max(releases$date)

# Filter row for last release
last_release <- filter(releases, date == last_release_date)

# Print last_release
last_release

# How long since last release?
Sys.Date() - last_release_date
```

### **`06-Getting datetimes into R`**

-   Use `as.POSIXct()` and an appropriate string to input the datetime corresponding to Oct 1st 2010 at 12:12:00.

-   Enter the same datetime again, but now specify the timezone as `"America/Los_Angeles"`.

-   Use `read_csv()` to read in `rversions.csv` again.

-   Examine the structure of the `datetime` column to verify `read_csv()` has correctly interpreted it as a datetime.

```{r}
# Use as.POSIXct to enter the datetime 
as.POSIXct("2010-10-01 12:12:00")

# Use as.POSIXct again but set the timezone to `"America/Los_Angeles"`
as.POSIXct("2010-10-01 12:12:00", tz = "America/Los_Angeles")

# Use read_csv to import rversions.csv
# already read
releases$datetime <- as.POSIXct(releases$datetime)


# Examine structure of datetime column
str(releases$datetime)
```

### **`07-Datetimes behave nicely too`**

-   Use `read_csv()` to import `cran-logs_2015-04-17.csv`.

-   Print `logs` to see the information we have on each download.

-   Store the R 3.2.0 release time as a `POSIXct` object.

-   Find out when the first request for 3.2.0 was made by filtering for values in the `datetime` column that are greater than `release_time`.

-   Finally see how downloads increase by creating histograms of download time for 3.2.0 and the previous version 3.1.3. We've provided most of the code, you just need to specify the `x` aesthetic to be the `datetime` column.

```{r}
# Use read_csv() to import rversions.csv
rversions_path <- file.path("..", "00_Datasets", "rversions.csv")
releases <- read.csv(rversions_path)
```

```{r}
# Import "cran-logs_2015-04-17.csv" with read_csv()
logs_path <- file.path("..", "00_Datasets", "cran-logs_2015-04-17.csv")

logs <- read_csv(logs_path)

# Print logs
print(logs)

# Store the release time as a POSIXct object
release_time <- as.POSIXct("2015-04-16 07:13:33", tz = "UTC")

# When is the first download of 3.2.0?
logs %>% 
  filter(datetime > release_time,
    r_version == "3.2.0")

# Examine histograms of downloads by version
ggplot(logs, aes(x = datetime)) +
  geom_histogram(bins = 30) +
  geom_vline(aes(xintercept = as.numeric(release_time)))+
  facet_wrap(~ r_version, ncol = 1)
```

### `The End`
