---
title: "Arithmetic with Dates and Times"
author: "Mohamad Osman"
date: '2022-06-29'
output: rmarkdown::github_document
---

# Section 03: Arithmetic with Dates and Times

### `01-How long has it been?`

To get finer control over a difference between datetimes use the `base` function [**`difftime()`**](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/difftime). For example instead of `time1 - time2`, you use `difftime(time1, time2)`.

[**`difftime()`**](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/difftime) takes an argument `units` which specifies the units for the difference. Your options are `"secs"`, `"mins"`, `"hours"`, `"days"`, or `"weeks"`.

To practice you'll find the time since the first man stepped on the moon. You'll also see the `lubridate` functions `today()` and [**`now()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/now) which when called with no arguments return the current date and time in your system's timezone.

-   Apollo 11 landed on July 20, 1969. Use `difftime()` to find the number of days between `today()` and `date_landing`.

-   Neil Armstrong stepped onto the surface at 02:56:15 UTC. Use `difftime()` to find the number of seconds between `now()` and `moment_step`.

```{r}
library(dplyr)
library(lubridate)
```

```{r}
# The date of landing and moment of step
date_landing <- mdy("July 20, 1969")
moment_step <- mdy_hms("July 20, 1969, 02:56:15", tz = "UTC")

# How many days since the first man on the moon?
difftime(today(), date_landing, units = "days")

# How many seconds since the first man on the moon?
difftime( now(), moment_step, units = "secs")

```

### **`02-How many seconds are in a day?`**

How many seconds are in a day? There are 24 hours in a day, 60 minutes in an hour, and 60 seconds in a minute, so there should be `24*60*60` = `86400` seconds, right?

Not always! In this exercise you'll see a counter example, can you figure out what is going on?

We've put code to define three times in your script - noon on March 11th, March 12th, and March 13th in 2017 in the US Pacific timezone.

-   Find the difference in time between `mar_13` and `mar_12` in seconds. *This should match your intuition*.

-   Now, find the difference in time between `mar_12` and `mar_11` in seconds. *Surprised?*

```{r}
# Three dates
mar_11 <- ymd_hms("2017-03-11 12:00:00", 
  tz = "America/Los_Angeles")
mar_12 <- ymd_hms("2017-03-12 12:00:00", 
  tz = "America/Los_Angeles")
mar_13 <- ymd_hms("2017-03-13 12:00:00", 
  tz = "America/Los_Angeles")

# Difference between mar_13 and mar_12 in seconds
difftime(mar_13, mar_12, units = "secs")

# Difference between mar_12 and mar_11 in seconds
difftime(mar_12, mar_11, units = "secs")

```

Good work. Why would a day only have 82800 seconds? At 2am on Mar 12th 2017, Daylight Savings started in the Pacific timezone. That means a whole hour of seconds gets skipped between noon on the 11th and noon on the 12th.

### **`03-Adding or subtracting a time span to a datetime`**

-   It's Monday Aug 27th 2018 at 2pm and you want to remind yourself this time next week to send an email. Add a period of one week to `mon_2pm`.

-   It's Tuesday Aug 28th 2018 at 9am and you are starting some code that usually takes about 81 hours to run. When will it finish? Add a duration of 81 hours to `tue_9am`.

-   What were you doing five years ago? Subtract a period of 5 years from `today()`.

-   Subtract a duration of 5 years from `today()`. *Will this give a different date?*

```{r}
# Add a period of one week to mon_2pm
mon_2pm <- dmy_hm("27 Aug 2018 14:00")
mon_2pm + ddays(7)

# Add a duration of 81 hours to tue_9am
tue_9am <- dmy_hm("28 Aug 2018 9:00")
tue_9am + dhours(81)

# Subtract a period of five years from today()
today() - years(x = 5)

# Subtract a duration of five years from today()
today() - dyears(5)
```

### **`04-Arithmetic with timespans`**

You can add and subtract timespans to create different length timespans, and even multiply them by numbers. For example, to create a duration of three days and three hours you could do: `ddays(3) + dhours(3)`, or `3*ddays(1) + 3*dhours(1)` or even `3*(ddays(1) + dhours(1))`.

There was an eclipse over North America on 2017-08-21 at 18:26:40. It's possible to predict the next eclipse with similar geometry by calculating the time and date one [**Saros**](https://eclipse.gsfc.nasa.gov/SEsaros/SEsaros.html) in the future. A Saros is a length of time that corresponds to 223 Synodic months, a Synodic month being the period of the Moon's phases, a duration of 29 days, 12 hours, 44 minutes and 3 seconds.

Do just that in this exercise

-   Create a duration corresponding to one Synodic Month: 29 days, 12 hours, 44 minutes and 3 seconds.

-   Create a duration corresponding to one Saros by multiplying `synodic` by 223.

-   Add `saros` to `eclipse_2017` to predict the next eclipse.

```{r}
# Time of North American Eclipse 2017
eclipse_2017 <- ymd_hms("2017-08-21 18:26:40")

# Duration of 29 days, 12 hours, 44 mins and 3 secs
synodic <- ddays(29) + dhours(12) + dminutes(44) + dseconds(3)

# 223 synodic months
saros <- 223 * synodic

# Add saros to eclipse_2017
saros + eclipse_2017
```

### **`05-Generating sequences of datetimes`**

By combining addition and multiplication with sequences you can generate sequences of datetimes. For example, you can generate a sequence of periods from 1 day up to 10 days with,

    1:10 * days(1)

Then by adding this sequence to a specific datetime, you can construct a sequence of datetimes from 1 day up to 10 days into the future

    today() + 1:10 * days(1)

-   Create `today_8am()` by adding a period of 8 hours to `today()`

-   Create a sequence of periods from one period of two weeks, up to 26 periods of two weeks.

-   Add `every_two_weeks` to `today_8am`.

```{r}
# Add a period of 8 hours to today
today_8am <- today() + hours(8)

# Sequence of two weeks from 1 to 26
every_two_weeks <- 1:26 * weeks(1)

# Create datetime for every two weeks for a year
today_8am + every_two_weeks

```

### **`06-The tricky thing about months`**

What should `ymd("2018-01-31") + months(1)` return? Should it be 30, 31 or 28 days in the future? Try it. In general `lubridate` returns the same *day of the month* in the next month, but since the 31st of February doesn't exist `lubridate` returns a missing value, `NA`.

There are alternative addition and subtraction operators: `%m+%` and `%m-%` that have different behavior. Rather than returning an `NA` for a non-existent date, they roll back to the last existing date.

You'll explore their behavior by trying to generate a sequence for the last day in every month this year.

We've put `jan_31`, the date for January 31st this year in your workspace.

-   Start by creating a sequence of 1 to 12 periods of 1 month.

-   Add `month_seq` to `jan_31`. *Notice what happens to any month where the 31st doesn't exist*

-   Now add `month_seq` to `jan_31` using the `%m+%` operator.

-   Try subtracting `month_seq` from `jan_31` using the `%m-%` operator.

```{r}
# A sequence of 1 to 12 periods of 1 month
month_seq <- 1:12 * months(1)
jan_31 <- as.Date("2022-01-31")
# Add 1 to 12 months to jan_31
month_seq + jan_31
print("--------------------------------------")

# Replace + with %m+%
month_seq %m+% jan_31
print("--------------------------------------")

# Replace + with %m-%
month_seq %m-% jan_31

```

### **`07-Examining intervals. Reigns of kings and queens`**

You can create an interval by using the operator `%--%` with two datetimes. For example `ymd("2001-01-01") %--% ymd("2001-12-31")` creates an interval for the year of 2001.

Once you have an interval you can find out certain properties like its start, end and length with `int_start()`, `int_end()` and `int_length()` respectively.

Practice by exploring the reigns of kings and queens of Britain (and its historical dominions).

We've put the data `monarchs` in your workspace.

-   Print `monarchs` to take a look at the data

-   Create a new column called `reign` that is an interval between `from` and `to`.

-   Create another new column, `length`, that is the interval length of `reign`. The rest of the pipeline we've filled in for you, it arranges by decreasing `length` and selects the `name`, `length` and `dominion` columns.

```{r}
library(readxl)


monarchs_file = file.path("..", "00_Datasets", "monarchs.xlsx")

monarchs <- read_excel(monarchs_file, sheet = 1, col_names = TRUE)
monarchs$from <- as.POSIXct(monarchs$from, format = "%Y-%m-%d %H:%M:%S")
monarchs$to <- as.POSIXct(monarchs$to, format = "%Y-%m-%d %H:%M:%S")
```

```{r}
# Print monarchs
 print(monarchs)

# Create an interval for reign
monarchs <- monarchs %>%
  mutate(reign = from %--% to) 

# Find the length of reign, and arrange
monarchs %>%
  mutate(length = int_length(reign)) %>% 
  arrange(desc(length)) %>%
  select(name, length, dominion)

```

### **`08-Comparing intervals and datetimes`**

The operator `%within%` tests if the datetime (or interval) on the left hand side is within the interval of the right hand side. For example, if `y2001` is the interval covering the year 2001,

    y2001 <- ymd("2001-01-01") %--% ymd("2001-12-31")

Then `ymd("2001-03-30") %within% y2001` will return `TRUE` and `ymd("2002-03-30") %within% y2001` will return `FALSE`.

`int_overlaps()` performs a similar test, but will return true if two intervals overlap at all.

We've put `halleys` a data set describing appearances of Halley's comet in your workspace.

-   Print `halleys` to examine the date. `perihelion_date` is the date the Comet is closest to the Sun. `start_date` and `end_date` are the range of dates the comet is visible from Earth.

-   Create a new column, `visible`, that is an interval from `start_date` to `end_date`.

-   You'll work with one appearance, extract the 14th row of `halleys.`

-   Filter `monarchs` to those where `halleys_1066$perihelion_date` is within `reign`.

-   Filter `monarchs` to those where `halleys_1066$visible` overlaps `reign`.

```{r}
halleys_file = file.path("..", "00_Datasets", "halleys.xlsx")

halleys <- read_excel(halleys_file, sheet = 1, col_names = TRUE)
halleys$perihelion_date <- as.Date(halleys$perihelion_date, format = "%Y-%m-%d")
halleys$start_date <- as.Date(halleys$start_date, format = "%Y-%m-%d")
halleys$end_date <- as.Date(halleys$end_date, format = "%Y-%m-%d")
halleys
```

```{r}
# Print halleys
print(halleys)

# New column for interval from start to end date
halleys <- halleys %>% 
  mutate(visible = start_date  %--% end_date)

# The visitation of 1066
halleys_1066 <- halleys[14, ] 

# Monarchs in power on perihelion date
monarchs %>% 
  filter(halleys_1066$perihelion_date %within% reign) %>%
  select(name, from, to, dominion)

# Monarchs whose reign overlaps visible time
monarchs %>% 
  filter(int_overlaps(halleys_1066$visible, reign)) %>%
  select(name, from, to, dominion)

```

**09-Converting to durations and periods**

Intervals are the most specific way to represent a span of time since they retain information about the exact start and end moments. They can be converted to periods and durations exactly: it's possible to calculate both the exact number of seconds elapsed between the start and end date, as well as the perceived change in clock time.

To do so you use the [**`as.period()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/as.period), and [**`as.duration()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/as.duration) functions, parsing in an interval as the only argument.

Try them out to get better representations of the length of the monarchs reigns.

-   Create new columns for `duration` and `period` that convert `reign` into the appropriate object.

-   Examine the `name`, `duration` and `period` columns.

```{r}
# New columns for duration and period
monarchs <- monarchs %>%
  mutate(
    duration = as.duration(reign),
    period = as.period(reign)) 
    
# Examine results    
monarchs %>%
  select(name, duration, period)
```

### `The End`

\

\

\

\

\

\

\
