---
title: "Problems in practice"
author: "Mohamad Osman"
date: '2022-07-04'
output: rmarkdown::github_document
---

# **Section 04: Problems in practice**

### **`01-Setting the timezone`**

If you import a datetime and it has the wrong timezone, you can set it with [**`force_tz()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/force_tz). Pass in the datetime as the first argument and the appropriate timezone to the `tzone` argument. Remember the timezone needs to be one from `OlsonNames()`.

I've put the times as listed on the FIFA website for games 2 and 3 in the group stage for New Zealand in your code.

-   Game 2 was played in Edmonton. Use `force_tz()` to set the timezone of game 2 to `"America/Edmonton"`.

-   Game 3 was played in Winnipeg. Use `force_tz()` to set the timezone of game 3 to `"America/Winnipeg"`.

-   Find out how long the team had to rest between the two games, by using `as.period()` on the interval between `game2_local` and `game3_local`.

```{r}
library(lubridate)
library(dplyr)
library(readr)

```

```{r}
# Game2: CAN vs NZL in Edmonton
game2 <- mdy_hm("June 11 2015 19:00")

# Game3: CHN vs NZL in Winnipeg
game3 <- mdy_hm("June 15 2015 18:30")

# Set the timezone to "America/Edmonton"
game2_local <- force_tz(game2, tzone = "America/Edmonton")
game2_local

# Set the timezone to "America/Winnipeg"
game3_local <- force_tz(game3, tzone = "America/Winnipeg")
game3_local

# How long does the team have to rest?
as.period(game2_local %--% game3_local)
```

Great work! Edmonton and Winnipeg are in different timezones, so even though the start times of the games only look 30 minutes apart, they are in fact 1 hour and 30 minutes apart, and the team only has 3 days, 22 hours and 30 minutes to prepare.

### **`02-Viewing in a timezone`**

To view a datetime in another timezone use [**`with_tz()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/with_tz). The syntax of [**`with_tz()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/with_tz) is the same as [**`force_tz()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/force_tz), passing a datetime and set the `tzone` argument to the desired timezone. Unlike [**`force_tz()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/force_tz), [**`with_tz()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/with_tz) isn't changing the underlying moment of time, just how it is displayed.

For example, the difference between `now()` displayed in the "America/Los_Angeles" timezone and "Pacific/Auckland" timezone is 0:

    now <- now()
    with_tz(now, "America/Los_Angeles") - 
      with_tz(now,  "Pacific/Auckland")

Help me figure out when to tune into the games from the previous exercise.

```{r}
# What time is game2_local in NZ?
with_tz(game2_local, tzone = "Pacific/Auckland")

# What time is game2_local in Corvallis, Oregon?
with_tz(game2_local, tzone = "America/Los_Angeles")

# What time is game3_local in NZ?
with_tz(game3_local, tzone = "Pacific/Auckland")
```

### **`03-Timezones in the weather data`**

```{r}
# Import "akl_weather_hourly_2016.csv"
akl_weather_h_path <- file.path("..", "00_Datasets", "akl_weather_hourly_2016.csv")
akl_hourly_raw <- read_csv(akl_weather_h_path)


# Use make_date() to combine year, month and mday 
akl_hourly  <- akl_hourly_raw  %>% 
  mutate(date = make_date(year = year, month = month, day = mday))

# Parse datetime_string 
akl_hourly <- akl_hourly  %>% 
  mutate(
    datetime_string = paste(date, time, sep = "T"),
    datetime = parse_date_time(datetime_string, orders = "ymdHMS")
  )

head(akl_hourly, 5)

```

The data is available in the `akl_hourly` data frame.

-   What timezone are `datetime` and `date_utc` currently in? Examine the head of the `datetime` and `date_utc` columns to find out.

-   Fix `datetime` to have the timezone for `"Pacific/Auckland"`.

-   Reexamine the head of the `datetime` column to check the times have the same clocktime, but are now in the right timezone.

-   Now tabulate up the difference between the `datetime` and `date_utc` columns. *It should be zero if our hypothesis was correct*.

```{r}
# Examine datetime and date_utc columns
head(akl_hourly$datetime)
head(akl_hourly$date_utc)
  
# Force datetime to Pacific/Auckland
akl_hourly <- akl_hourly %>%
  mutate(
    datetime = force_tz(datetime, tzone = "Pacific/Auckland"))

# Reexamine datetime
head(akl_hourly$datetime)
  
# Are datetime and date_utc the same moments
table(akl_hourly$datetime - akl_hourly$date_utc)
  
```

### **`04-Times without dates`**

-   Use `read_csv()` to read in `"akl_weather_hourly_2016.csv"`. *`readr` knows about the `hms` class, so if it comes across something that looks like a time it will use it.*

-   In this case the `time` column has been parsed as a time without a date. Take a look at the structure of the `time` column to verify it has the class `hms`.

-   `hms` objects print like times should. Take a look by examining the head of the `time` column.

-   You can use `hms` objects in plots too. Create a plot with `time` on the x-axis, `temperature` on the y-axis, with lines grouped by `date`.

```{r}
library(tidyverse)

# Import auckland hourly data 
akl_hourly <- read_csv(akl_weather_h_path)

# Examine structure of time column
str(akl_hourly$time)

# Examine head of time column
head(akl_hourly$time)

# A plot using just time
ggplot(akl_hourly, aes(x = time, y = temperature)) +
  geom_line(aes(group = make_date(year, month, mday)), alpha = 0.2)

```

### `05-Fast parsing with fasttime`

The `fasttime` package provides a single function [**`fastPOSIXct()`**](https://www.rdocumentation.org/packages/fasttime/versions/1.0-2/topics/fastPOSIXct), designed to read in datetimes formatted according to ISO 8601. Because it only reads in one format, and doesn't have to guess a format, it is really fast!

You'll see how fast in this exercise by comparing how fast it reads in the dates from the Auckland hourly weather data (over 17,000 dates) to `lubridate`s [**`ymd_hms()`**](https://www.rdocumentation.org/packages/lubridate/versions/1.8.0/topics/ymd_hms).

To compare run times you'll use the [**`microbenchmark()`**](https://www.rdocumentation.org/packages/microbenchmark/versions/1.4.9/topics/microbenchmark) function from the package of the same name. You pass in as many arguments as you want each being an expression to time.

We've loaded the datetimes from the Auckland hourly data as strings into the vector `dates`.

-   Examine the structure of `dates` to verify it is a string and in the ISO 8601 format.

-   Parse `dates` with `fasttime` and pipe to `str()` to verify `fastPOSIXct` parses them correctly.

-   Now to compare timing, call `microbenchmark` where the first argument uses `ymd_hms()` to parse `dates` and the second uses `fastPOSIXct()`.

```{r}
dates <- c("2015-12-31T11:00:00Z", "2015-12-31T11:30:00Z", "2015-12-31T12:00:00Z",
"2015-12-31T12:30:00Z", "2015-12-31T13:00:00Z", "2015-12-31T13:30:00Z",
"2015-12-31T14:00:00Z", "2015-12-31T14:30:00Z", "2015-12-31T15:00:00Z",
"2015-12-31T15:30:00Z", "2015-12-31T16:00:00Z", "2015-12-31T16:30:00Z",
"2015-12-31T17:00:00Z", "2015-12-31T17:30:00Z", "2015-12-31T18:00:00Z",
"2015-12-31T18:30:00Z", "2015-12-31T19:00:00Z", "2015-12-31T19:30:00Z",
"2015-12-31T20:00:00Z", "2015-12-31T20:30:00Z", "2015-12-31T21:00:00Z",
"2015-12-31T21:30:00Z", "2015-12-31T22:00:00Z", "2015-12-31T22:30:00Z",
"2015-12-31T23:00:00Z", "2015-12-31T23:30:00Z", "2016-01-01T00:00:00Z",
"2016-01-01T00:30:00Z", "2016-01-01T01:00:00Z", "2016-01-01T01:30:00Z",
"2016-01-01T02:00:00Z", "2016-01-01T02:30:00Z", "2016-01-01T03:00:00Z",
"2016-01-01T03:30:00Z", "2016-01-01T04:00:00Z", "2016-01-01T04:30:00Z",
"2016-01-01T05:00:00Z", "2016-01-01T05:30:00Z", "2016-01-01T06:00:00Z",
"2016-01-01T06:30:00Z", "2016-01-01T07:00:00Z", "2016-01-01T07:30:00Z",
"2016-01-01T08:00:00Z", "2016-01-01T08:30:00Z", "2016-01-01T09:00:00Z",
"2016-01-01T09:30:00Z", "2016-01-01T10:00:00Z", "2016-01-01T10:30:00Z", 
"2016-01-01T11:00:00Z", "2016-01-01T11:30:00Z")
```

```{r}
library(microbenchmark)
library(fasttime)

# Examine structure of dates
str(dates)

# Use fastPOSIXct() to parse dates
fastPOSIXct(dates) %>% str(dates)

# Compare speed of fastPOSIXct() to ymd_hms()
microbenchmark(
  ymd_hms = ymd_hms(dates),
  fasttime = fastPOSIXct(dates),
  times = 20)
```

Original Results:

`Unit: microseconds      expr       min         lq      mean    median        uq        max neval   ymd_hms 28883.894 39871.5555 48564.648 43363.405 46950.459 128872.221    20  fasttime   708.559   925.7805  1152.037  1230.428  1329.998   1663.331    20`{=}

Great job! To compare speed, you can compare the average run time in the mean column. You should see `fasttime` is about 20 times faster than `ymd_hms()`

### 
**06-Fast parsing with lubridate::fast_strptime**

`dates` is in your workspace again.

-   Examine the head of `dates`. *What components are present? What separators are used?*

-   Parse `dates` with `fast_strptime()` by specifying the appropriate format string.

-   Compare the timing of `fast_strptime()` to `fasttime` and `ymd_hms()`.

```{r}
# Head of dates
head(dates)

# Parse dates with fast_strptime
fast_strptime(dates, 
    format = "%Y-%m-%dT%H:%M:%SZ") %>% str()

# Comparse speed to ymd_hms() and fasttime
microbenchmark(
  ymd_hms = ymd_hms(dates),
  fasttime = fastPOSIXct(dates),
  fast_strptime = fast_strptime(dates, 
    format = "%Y-%m-%dT%H:%M:%SZ"),
  times = 20)
```

### **07-Outputting pretty dates and times**

-   Create a `stamp()` based on the string `"Saturday, Jan 1, 2000"`.

-   Print `date_stamp`. *Notice it is a function.*

-   Pass `today()` to `date_stamp` to format today's date.

-   Now output today's date in American style `MM/DD/YYYY`.

-   Finally, use stamp based on the `finished` string I've put in your workspace to format `today()`.

```{r}
finished <- "I finished 'Dates and Times in R' on Thursday, September 4, 2017!"

# Create a stamp based on "Saturday, Jan 1, 2000"
date_stamp <- stamp("Saturday, Jan 1, 2000")

# Print date_stamp
print(date_stamp)

# Call date_stamp on today()
date_stamp(today())

# Create and call a stamp based on "12/31/1999"
stamp("12/31/1999")(today())

# Use string finished for stamp()
stamp(finished)(today())
```

### `The End`

\

\
