---
title: "How to write a function"
author: "Mohamad Osman"
date: '2022-07-05'
output: rmarkdown::github_document
---

# **Section 01: How to write a function**

### **`01-Calling functions`**

*The final line calculates the median number of gold medals each country won.*

Rewrite the call to [**`median()`**](https://www.rdocumentation.org/packages/stats/topics/median), following best practices.

```{r}
library(dplyr)
library(tidyverse)
```

```{r}
names <- c("USA", "GBR", "CHN", "RUS", "GER", "JPN", "FRA", "KOR", "ITA", "AUS", "NED", "HUN", "BRA", "ESP", "KEN", "JAM", "CRO", "CUB", "NZL", "CAN", "UZB", "KAZ", "COL", "SUI", "IRI", "GRE", "ARG", "DEN", "SWE", "RSA", "UKR", "SRB", "POL", "PRK", "BEL", "THA", "SVK", "GEO", "AZE", "BLR", "TUR", "ARM", "CZE", "ETH", "SLO", "INA", "ROU", "BRN", "VIE", "TPE", "BAH", "IOA", "CIV", "FIJ", "JOR", "KOS", "PUR", "SIN", "TJK", "MAS", "MEX", "VEN", "ALG", "IRL", "LTU", "BUL", "IND", "MGL", "BDI", "GRN", "NIG", "PHI", "QAT", "NOR", "EGY", "TUN", "ISR", "AUT", "DOM", "EST", "FIN", "MAR", "NGR", "POR", "TTO", "UAE", "IOC")

gold_medals <- c(46, 27, 26, 19, 17, 12, 10, 9, 8, 8, 8, 8, 7, 7, 6, 6, 5, 5, 4, 4, 4, 3,3, 3, 3, 3, 3, 2, 2,  2, 2, 2, 2, 2, 2, 2, 2, 2, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, NA)
#names(gold_medals) <- attr

names(gold_medals) <- names

typeof(gold_medals)

```

```{r}
# Look at the gold medals data
gold_medals

# Note the arguments to median()
args(median)

# Rewrite this function call, following best practices
median(gold_medals, na.rm = TRUE)
```

*The final line calculates each country's ranking by number of gold medals. It uses negative `gold_medals` so that the country with the most medals will have 1st place: the largest positive value in `gold_medals` is the smallest ("most negative") value in `-gold_medals`.*

Rewrite the call to [**`rank()`**](https://www.rdocumentation.org/packages/base/topics/rank), following best practices.

```{r}
# Note the arguments to rank()
args(rank)

# Rewrite this function call, following best practices
rank(-gold_medals, na.last = "keep", ties.method = "min")
```

### **`02-Your first function: tossing a coin`**

Your first function: tossing a coin

```{r}
coin_sides <- c("head", "tail")

# Sample from coin_sides once
sample(coin_sides, size = 1)
```

-   Write a template for your function, naming it `toss_coin`. The function should take no arguments. Don't include the body of the function yet.

-   Copy your script, and paste it into the function body.

-   Call your function.

```{r}
# Write a template for your function, toss_coin()
toss_coin <- function() {
  # (Leave the contents of the body for later)
# Add punctuation to finish the body
  sample(coin_sides, size = 1)
} 

toss_coin()
```

### **`03-Inputs to functions`**

-   Sample from `coin_sides` `n_flips` times with replacement.

```{r}
coin_sides <- c("head", "tail")
n_flips <- 10

# Sample from coin_sides n_flips times with replacement
sample(coin_sides, size = n_flips, replace =TRUE)
```

-   Update the definition of `toss_coin()` to accept a single argument, `n_flips`. The function should sample `coin_sides` `n_flips` times with replacement. *Remember to change the signature and the body.*

-   Generate 10 coin flips.

```{r}
# Update the function to return n coin tosses
toss_coin <- function(n_flips = 1) {
  coin_sides <- c("head", "tail")
  sample(coin_sides, n_flips, replace = TRUE)
}

# Generate 10 coin tosses
toss_coin(n_flips = 10)
```

### **`04-Multiple inputs to functions`**

-   Bias the coin by weighting the sampling. Specify the `prob` argument so that heads are sampled with probability `p_head` (and tails are sampled with probability `1 - p_head`).

```{r}
coin_sides <- c("head", "tail")
n_flips <- 10
p_head <- 0.8

# Define a vector of weights
weights <- c(p_head, 1 - p_head)

# Update so that heads are sampled with prob p_head
sample(coin_sides, n_flips, replace = TRUE, prob = weights)
```

-   Update the definition of `toss_coin()` so it accepts an argument, `p_head`, and weights the samples using the code you wrote in the previous step.

-   Generate 10 coin tosses with an 80% chance of each head.

```{r}
# Update the function so heads have probability p_head
toss_coin <- function(n_flips, p_head) {
  coin_sides <- c("head", "tail")
  # Define a vector of weights
  weights <- c(p_head, 1-p_head)
  # Modify the sampling to be weighted
  sample(coin_sides, n_flips, replace = TRUE, prob = weights)
}

# Generate 10 coin tosses
toss_coin(10, 0.8)
```

### **`05-Renaming GLM`**

-   Run a generalized linear regression by calling `glm()`. Model `n_visits` vs. `gender`, `income`, and `travel` on the `snake_river_visits` dataset, setting the `family` to `poisson`.

```{r}
# Load data
file_path <- file.path("..", "00_Datasets", "snake_river_visits.rds")
snake_river_visits <- readRDS(file_path)

# Run a generalized linear regression 
glm(
  # Model no. of visits vs. gender, income, travel
  n_visits ~ gender + income + travel, 
  # Use the snake_river_visits dataset
  data = snake_river_visits, 
  # Make it a Poisson regression
  family = poisson
)

```

-   Define a function, `run_poisson_regression()`, to run a Poisson regression. This should take two arguments: `data` and `formula`, and call `glm()`, passing those arguments and setting `family` to `poisson`.

```{r}
# Write a function to run a Poisson regression
run_poisson_regression <- function(data, formula) {

    glm(formula, data, family = poisson)
}
```

-   Recreate the Poisson regression model from the first step, this time by calling your `run_poisson_regression()` function.

```{r}
snake_river_explanatory <- snake_river_visits

# From previous step
run_poisson_regression <- function(data, formula) {
  glm(formula, data, family = poisson)
}

# Re-run the Poisson regression, using your function
model <- snake_river_visits %>%
  run_poisson_regression(n_visits ~ gender + income + travel)

# Run this to see the predictions
snake_river_explanatory %>%
  mutate(predicted_n_visits = predict(model, ., type = "response"))%>%
  arrange(desc(predicted_n_visits))
```

### `The End` 

\



\

\

\

\

\

\
